### 2023-08-15 paper_codes.Rmd
Add files via upload

---
title: "Paper Codes"
author: "Markus Viljanen"
date: "2023-08-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(plyr)
library(tidyverse)
library(sf)
library(INLA)
library(mgcv)
library(ggplot2)
library(cowplot)
library(colorspace)
library(scales)
theme_set(theme_bw())
```

## Data set

Species to analyze and their Latin names

```{r}
species.map <- read.csv(file="data_paper/chosen_species.csv") %>% transmute(species=Nednaam, name=Wetnaam)
species <- species.map$species
```

Netherlands boundary shapefile

```{r}
NL <- st_read("data_paper/NL.shp")$geometry#st_simplify()
NL <- (NL - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
```

Data for vegetation plots and abiotic measurements data

```{r}
# Load field visit data
df.plots <- read.csv(file="data_paper/data_plots.csv", check.names = F)
# Add geometries
sf.plots <- df.plots %>% st_as_sf(coords = c("X","Y"), crs=st_crs(NL))
sf.plots[c("X", "Y")] <- sf.plots  %>% st_coordinates()
```

Data for Netherlands grid

```{r}
# Load Netherlands grid data
df.grid <- read.csv(file="data_paper/data_grid.csv", check.names = F)
# Grid with rectangles and their centers
grid.rect <- st_make_grid(NL,n=100,crs=st_crs(NL), what='polygons', square=T) 
grid.circle <- st_make_grid(NL,n=100,crs=st_crs(NL), what='centers')
index <- which(lengths(st_within(grid.circle, NL)) > 0) #st_intersects(grid.rect, NL)
# Add geometries like originally created
geoms <- data.frame(grid.circle[index], grid.rect[index]) 
sf.grid <- cbind(df.grid, geoms) %>% st_as_sf(crs=st_crs(NL))
```

## Plot field visits

Translate landuse to English

```{r}
sf.grid$landuse <- recode(sf.grid$bodemgebruik, "Bos"="Forest", "Bebouwd"="Built", 
                          "Droog natuurlijk terrein"="Natural terrain (dry)",
                          "Hoofdweg"="Highway","Landbouw en overig agrarisch"="Agricultural", 
                          "Nat natuurlijk terrein"= "Natural terrain (wet)", 
                          "Recreatie"="Recreational", "Spoorterrein"="Railway", "Water"="Water")
sf.grid$landuse <- fct_relevel(sf.grid$landuse, "Forest", "Natural terrain (dry)", "Natural terrain (wet)", 
                               "Agricultural", "Recreational", "Highway", "Railway", "Water", "Built")
sf.plots$landuse <- recode(sf.plots$bodemgebruik, "Bos"="Forest", "Bebouwd"="Built", 
                          "Droog natuurlijk terrein"="Natural terrain (dry)",
                          "Hoofdweg"="Highway","Landbouw en overig agrarisch"="Agricultural", 
                          "Nat natuurlijk terrein"= "Natural terrain (wet)", 
                          "Recreatie"="Recreational", "Spoorterrein"="Railway", "Water"="Water")
sf.plots$landuse <- fct_relevel(sf.plots$landuse, "Forest", "Natural terrain (dry)", "Natural terrain (wet)", 
                               "Agricultural", "Recreational", "Highway", "Railway", "Water", "Built")
```

Plot landuse in each grid cell:

```{r}
ggplot(NL) + geom_sf() +
	geom_sf(aes(geometry=geometry.1, fill=landuse), data=sf.grid) +
  scale_fill_manual(values=c("Forest"="forestgreen", "Built"="slategrey",
                             "Natural terrain (dry)"="peru","Natural terrain (wet)"="lightseagreen",
                             "Highway"="grey","Railway"="seashell2", "Agricultural"="wheat",#
                             "Recreatie"="lawngreen", "Water"="dodgerblue"
                             )) +
	theme(legend.position="right") #+ggtitle('Grid of Netherlands: Landuse') 
ggsave('results_paper/landuse.png', width=1600, heigh=1200, units='px')
```

Bar plot of landuse in grid cells vs. vegetation plots:

```{r}
f.grid <- table(sf.grid$landuse)
f.plots <- table(sf.plots$landuse)
f <- rbind(data.frame(f.grid / sum(f.grid), Data='Grid'),
           data.frame(f.plots / sum(f.plots), Data='Plots'))

ggplot(data=f, aes(x=Var1, y=Freq, fill=Data)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual(values=c('#999999','#E69F00')) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab('Landuse') + ylab('% of total') #+ ggtitle('Landuse: Grid of Netherlands vs. Field visits')
ggsave('results_paper/landuse_vs_visit.png', width=1400, heigh=900, units='px')
```

Plot field visits to vegetation plots and abiotic measurements over the grid:

```{r}
sf.plots$is_fieldvisit <- !is.na(sf.plots$Dagnummer)
sf.plots$has_abiotics <- !is.na(sf.plots$releve_nr)
```

```{r}
ggplot(NL) + geom_sf() +
	geom_sf(aes(geometry=geometry.1), data=sf.grid) + 
	#geom_sf(aes(geometry=geometry), shape=3, size=0.3, data=sf.grid) + 
	geom_sf(aes(geometry=geometry, color=has_abiotics), shape=4, size=0.5, data=sf.plots) + 
	theme(plot.title = element_text(hjust = 0.5), legend.title=element_text(size=10), 
		legend.text=element_text(size=8), legend.position="right") +
	scale_colour_manual(values=c("#377EB8", "#4DAF4A"), name="Has abiotics") #+ 
	#ggtitle('Location of Field Visits and Abiotics') 
ggsave('results_paper/visits.png', width=1400, heigh=1200, units='px')

```

## Spatial interpolation of common heather

Create a mesh over the Netherlands for spatial autocorrelation modelling

```{r}
step <- 10000 / 1000
domain <- inla.nonconvex.hull(as.matrix(sf.grid[c("X","Y")])[,1:2], convex=-0.04)
mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step, step*4), cutoff=step/4)
```

Illustrate the mesh and common heather (Struikhei) occurrence

```{r}

plot(NL, col="grey", main='Gaussian Markov Random Field')
plot(mesh, asp=1, add=T)
rbPal <- colorRampPalette(c('lightyellow','green'))
col <- rbPal(2)[sf.plots[["Struikhei"]] + 1]
plot(sf.plots["Struikhei"], pch=21, bg=col, pal=rbPal, cex=0.2, add=T, alpha=0.2)
legend(-200,150, text.font=4, cex=0.6, legend=c("Not Observed", "Observed"), col=c("lightyellow", "green"), pch=21)#16000, 614000
```

```{r}
## Prior: log(range) ~N(0, 10), log(sigma) = N(0,10)
#range0 <- 1 #exp(10) # ~22 km
#sigma0 <- 1 # ~2.7
#kappa0 <- sqrt(8)/range0
#tau0 <- 1/(sqrt(4*pi)*kappa0*sigma0) #should be 8?
#spde <- inla.spde2.matern(mesh,alpha=2,
#                          B.tau=matrix(c(log(tau0), -1, +1), nrow=1, ncol=3),
#                          B.kappa=matrix(c(log(kappa0), 0, -1), nrow=1, ncol=3),
#                          theta.prior.mean=c(0,0), theta.prior.prec=c(0.1,0.1))
#s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde) 
```

Interpolate common heather occurrence in the Netherlands with simple kriging, save zscore of log-odds as pH indicator:

```{r}
spde <- inla.spde2.matern(mesh=mesh,alpha=2)
s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde)

A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.plots %>% as.data.frame %>% select(X,Y)))
A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.grid %>% as.data.frame %>% select(X,Y)))

stack.est <- inla.stack(data=list(y = sf.plots[["Struikhei"]]),
                        A=list(A_y), effects=list(c(s.index, list(intercept=1))), 
                        tag="est")
stack.pred <- inla.stack(data=list(y=NA), 
                         A=list(A_pred), effects=list(c(s.index, list(intercept=1))),
                         tag="pred") 
joint.stack <- inla.stack(stack.est, stack.pred)

formula <- y ~ -1 + intercept + f(spatial.field, model=spde)

output.stack <- inla(formula, family="binomial", 
               data=inla.stack.data(joint.stack, spde=spde),
               control.predictor=list(A=inla.stack.A(joint.stack), compute=TRUE, link=1))

# Save extrapolated Struikhei as indicator of pH
plot.indicator <- output.stack$summary.linear.predictor[inla.stack.index(joint.stack, tag="est")$data,"mean"]
grid.indicator <- output.stack$summary.linear.predictor[inla.stack.index(joint.stack, tag="pred")$data,"mean"]
sf.plots$pHindicator <- (plot.indicator - mean(grid.indicator))/sd(grid.indicator)
sf.grid$pHindicator <- (grid.indicator - mean(grid.indicator))/sd(grid.indicator)
```

Plot observed vs. predicted common heather prevalence

```{r}
# Calculate observed Strukhei at each grid point
results <- sf.grid %>% st_set_geometry("geometry.1") %>% 
  st_join(sf.plots%>%select(`Struikhei`), st_contains) %>% 
  group_by(geometry.1) %>% summarize(Struikhei=mean(`Struikhei`, na.rm=T))
# Save predicted Strukhei at each grid point
index.pred.response <- inla.stack.index(joint.stack, tag="pred")$data
results$Struikhei.mean <- output.stack$summary.fitted.values[index.pred.response, "mean"]
results$Struikhei.sd <- output.stack$summary.fitted.values[index.pred.response,"sd"]

plot1 <- ggplot(NL) + geom_sf() +
	geom_sf(aes(geometry=geometry.1, fill=Struikhei), data=results) + 
  scale_fill_gradient(low="lightyellow", high="green", limits=c(0,1))+
	ggtitle('Observed Struikhei') + theme(legend.position="bottom")

plot2 <- ggplot(NL) + geom_sf() +
	geom_sf(aes(geometry=geometry.1, fill=Struikhei.mean), data=results) + 
  scale_fill_gradient(low="lightyellow", high="green", limits=c(0,1))+
	ggtitle('Predicted Struikhei') + theme(legend.position="bottom")

plot_grid(plot1, plot2, nrow = 1)
ggsave('results_paper/Struikhei.png', width=2000, heigh=1400, units='px')
```

## Spatial interpolation of pH

Illustrate the mesh and pH measurements:

```{r}
plot(NL, col="grey", main='Gaussian Markov Random Field')
plot(mesh, asp=1, add=T)
rbPal <- colorRampPalette(c('blue','red'))
col <- rbPal(10)[as.numeric(cut(sf.plots[["pH-H20"]],breaks = 10))]
plot(sf.plots["pH-H20"], pch=21, bg=col, pal=rbPal, cex=0.5, add=T)
legend(-200,150, text.font=4, cex=0.6, legend=c("Low", "High"), col=c("blue", "red"), pch=21)#16000, 614000
```

Interpolate pH in the Netherlands with kriging using soil type, landuse, and indicator plant (common heather) as covariates

```{r}
A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.plots %>% as.data.frame %>% select(X,Y)))
A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.grid %>% as.data.frame %>% select(X,Y)))

stack.est <- inla.stack(data=list(y = sf.plots[["pH-H20"]]),
                        A=list(A_y,1,1,1), effects=list(c(s.index, list(intercept=1)),
                                                  list(OMSCHRIJVI = sf.plots$OMSCHRIJVI),
                                                  list(bodemgebruik = sf.plots$bodemgebruik),
                                                  list(pHindicator = sf.plots$pHindicator)), 
                        tag="est")
stack.pred <- inla.stack(data=list(y=NA), 
                         A=list(A_pred,1,1,1), effects=list(c(s.index, list(intercept=1)),
                                                          list(OMSCHRIJVI = sf.grid$OMSCHRIJVI),
                                                          list(bodemgebruik = sf.grid$bodemgebruik),
                                                          list(pHindicator = sf.grid$pHindicator)),
                         tag="pred") 
joint.stack <- inla.stack(stack.est, stack.pred)

formula <- y ~ -1 + intercept + OMSCHRIJVI + bodemgebruik + pHindicator +f(spatial.field, model=spde) #+bodemgebruik

output.stack <- inla(formula, family="gaussian", 
               data=inla.stack.data(joint.stack, spde=spde),
               #control.fixed = list(expand.factor.strategy = "inla"),
               control.predictor=list(A=inla.stack.A(joint.stack), compute=TRUE))

output.field <- inla.spde2.result(inla=output.stack, name="spatial.field", spde=spde, do.transf=TRUE)
```

Illustrate the model parameters: intercept, measurement error, spatial variance and spatial range

```{r}
### Plot posterior distributions
par(mfrow=c(2,2))
# Plot for beta0
xrange <- range(output.stack$marginals.fix[[1]][,1])
yrange <- range(output.stack$marginals.fix[[1]][,2])
plot(output.stack$marginals.fix[[1]], type='l', xlim=xrange, ylim=yrange, xlab=expression(b[0]),ylab="")

# Plot for the variance sigma2e
s2e.marg <- inla.tmarginal(function(x) 1/x, output.stack$marginals.hy[[1]])
xrange <- range(s2e.marg[,1])
yrange <- range(s2e.marg[,2])
plot.default(s2e.marg, type='l', xlim=xrange, ylim=yrange, xlab=expression(sigma[e]^2),ylab="")

# Plot for the variance sigma2
xrange <- range(output.field$marginals.variance.nominal[[1]][,1])
yrange <- range(output.field$marginals.variance.nominal[[1]][,2])
plot(output.field$marginals.variance.nominal[[1]], type='l', xlim=xrange, ylim=yrange, xlab=expression(sigma^2),ylab="")

#Plot for the range
xrange <- range(output.field$marginals.range.nominal[[1]][,1])
yrange <- range(output.field$marginals.range.nominal[[1]][,2])
plot(output.field$marginals.range.nominal[[1]], type='l', xlim=xrange, ylim=yrange, xlab=expression(r),ylab="")

```

Plot observed vs. predicted pH

```{r}
# Calculate observed pH at each grid point
results <- sf.grid %>% st_set_geometry("geometry.1") %>% 
  st_join(sf.plots%>%select(`pH-H20`), st_contains) %>% 
  group_by(geometry.1) %>% summarize(pH=mean(`pH-H20`, na.rm=T))
# Save predicted pH at each grid point
index.pred.response <- inla.stack.index(joint.stack, tag="pred")$data
results$pH.mean <- output.stack$summary.fitted.values[index.pred.response,"mean"]
results$pH.sd <- output.stack$summary.fitted.values[index.pred.response,"sd"]

# Plot (ggplot)
plot1 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=pH), data=results) + 
  scale_fill_gradient2(low='red', midpoint=7, high='blue') + ggtitle('Observed pH') + theme(legend.position="bottom")
plot2 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=pH.mean), data=results) + 
  scale_fill_gradient2(low='red', midpoint=7, high='blue') + ggtitle('Predicted pH') + theme(legend.position="bottom")
#plot_grid(plot1, plot2, nrow = 1) + theme(plot.background = element_rect(fill = "white", colour = NA)) #, plot3
#ggsave('results_paper/pH.png', width=2000, heigh=1400, units='px')

plot.all <- ggdraw() +
  draw_plot(plot1, x = 0, y = 0, width = 0.50, height = 0.95) +
  draw_plot(plot2, x = 0.50, y = 0, width = 0.50, height = 0.95) +
  draw_plot_label(label = c("A", "B"), size = 15, x = c(0.01, 0.50), y = c(0.99,0.99)) +
  theme(plot.background = element_rect(fill = "white", colour = NA))
plot.all
ggsave('results_paper/pH.png', width=2000, heigh=1400, units='px')
```

Same plot with a third panel visualizing uncertainty of interpolated pH

```{r}
plot3 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=pH.sd), data=results) + 
  scale_fill_gradient(low='white', high='red', limits=c(0,1.11)) + ggtitle('Uncertainty pH') + theme(legend.position="bottom")
plot_grid(plot1, plot2, plot3, nrow = 1) + theme(plot.background = element_rect(fill = "white", colour = NA)) #, plot3
ggsave('results_paper/pH2.png', width=3000, heigh=1400, units='px')
```

Save interpolated pH for modelling

```{r}
index.pred.response <- inla.stack.index(joint.stack, tag="est")$data
sf.plots$pH <- output.stack$summary.fitted.values[index.pred.response,"mean"]
index.pred.response <- inla.stack.index(joint.stack, tag="pred")$data
sf.grid$pH <- output.stack$summary.fitted.values[index.pred.response,"mean"]
```

Histogram of pH in grid cells vs. vegetation plots

```{r}
f <- rbind(data.frame(pH = sf.grid$pH, Data='Grid'),
           data.frame(pH = sf.plots$pH, Data='Plots'))

ggplot(f, aes(x=pH, fill=Data)) +
  geom_histogram(aes(y=after_stat(density)), bins=50, position="dodge") +
  scale_fill_manual(values=c('#999999','#E69F00'))

```

## Process data

Marginal distribution statistics:

```{r}
n <- sum(sf.plots$is_fieldvisit)
sf.plots %>% as.data.frame %>% transmute(
  is_fieldvisit, has_abiotics, Dagnummer = Dagnummer, 
  pH = pH, ORG_STOF = log(ORG_STOF), CN = log(`C/N`), 
  N_totaal = log(N_totaal), P_totaal = log(P_totaal),
  K_CaCl2 = log(ifelse(K_CaCl2<=0, 0.1, K_CaCl2))) %>%
  summarize(across(all_of(c('Dagnummer', 'pH', 'ORG_STOF', 'CN','N_totaal', 'P_totaal', 'K_CaCl2')), 
                   list(mean=function(x) mean(x,na.rm=T), sd=function(x) sd(x,na.rm=T),
                        available = function(x) sum(!is.na(x) & is_fieldvisit)/n))) %>% round(2)
```
Z-score abiotic variables and clip extreme outliers to ensure equal impact of priors

```{r}
to_zscore <- function(x) (x - mean(x, na.rm=T))/sd(x, na.rm=T)
zclip <- function(x) ifelse(x > 3, 3, ifelse(x < -3, -3, x))

pH.mean = 7 #mean(sf.plots$pH)
pH.sd = 1 #sd(sf.plots$pH)

sf.plots <- sf.plots %>% mutate(
 Pq_grootte=replace_na(Pq_grootte, 25),
 Dagnummer = replace_na(Dagnummer, 200),
 Dagnummer.z = (Dagnummer-200)/30.44, 
 pH.z = zclip((pH - pH.mean)/pH.sd), 
 ORG_STOF.z = zclip(to_zscore(log(ORG_STOF))), 
 CN.z = zclip(to_zscore(log(`C/N`))),
 N_totaal.z = zclip(to_zscore(log(N_totaal))), 
 P_totaal.z = zclip(to_zscore(log(P_totaal))), 
 K_CaCl2.z = zclip(to_zscore(log(ifelse(K_CaCl2<=0, 0.1, K_CaCl2)))) 
)

sf.grid <- sf.grid %>% mutate(
 Dagnummer=200, Pq_grootte=25,
 Dagnummer.z = (Dagnummer-200)/30.44, 
 pH.z =  zclip((pH - pH.mean)/pH.sd)
)

```

Landuse coding for the spline

```{r}
code_terrain <- function (x) switch(x, "Bos"=-4, "Droog natuurlijk terrein"=-3, "Nat natuurlijk terrein"=-2, "Water"=-1,
                                     "Recreatie"=0, "Spoorterrein"=1, "Hoofdweg"=2,  "Bebouwd"=3, "Landbouw en overig agrarisch"=4)
sf.plots$terrain <- sapply(sf.plots$bodemgebruik %>% as.character, code_terrain)
sf.grid$terrain <- sapply(sf.grid$bodemgebruik %>% as.character, code_terrain)
```

Matching coordinate matrix

```{r}
xy.plots <- sf.plots %>% as.data.frame %>% select(X,Y)
xy.grid <- sf.grid %>% as.data.frame %>% select(X,Y)
```

For the cluster use these pre-processed files

```{r}
sf.plots %>% saveRDS("sf.plots.RDS")
sf.grid %>% saveRDS("sf.grid.RDS")
```

## Experiments: fit both models to data

Uniform mesh for the spatial component

```{r}
step <- 10000 / 1000
domain <- inla.nonconvex.hull(as.matrix(xy.plots), convex=-0.04)
mesh <- inla.mesh.2d(boundary=domain, max.edge=c(step*2, step*4), cutoff=step)
```

```{r}
spde <- inla.spde2.matern(mesh=mesh,alpha=2)##print(spde$n.spde)
s.index <- inla.spde.make.index(name="spatial.field", n.spde=spde$n.spde)

A_y <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.plots))
A_pred <- inla.spde.make.A(mesh=mesh, loc=as.matrix(xy.grid))
```

### Two-stage model

Predict missing abiotic variables

```{r}
terms.abiotics <- list()
i <- 1
for (variable in c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")) {
  print(variable)

  stack.est <- inla.stack(data=list(y = sf.plots[[variable]]),
                          A=list(A_y), effects=list(c(s.index, list(intercept=1))), 
                          tag="est")
  stack.pred <- inla.stack(data=list(y=NA), 
                           A=list(A_pred), effects=list(c(s.index, list(intercept=1))),
                           tag="pred") 
  join.stack <- inla.stack(stack.est, stack.pred)

  formula <- y ~ -1 + intercept + f(spatial.field, model=spde)

  output <- inla(formula, family="gaussian", 
                      data=inla.stack.data(join.stack, spde=spde),
                      control.predictor=list(A=inla.stack.A(join.stack), compute=TRUE, link=1))

  # predictions at field visits
  index.est <- inla.stack.index(join.stack, tag="est")$data
  variable.est <- output$summary.fitted.values[index.est,]
  colnames(variable.est) <- sprintf("%s.%s", variable, colnames(variable.est))
  sf.plots[colnames(variable.est)] <- variable.est

  # predictions at grid
  index.pred <- inla.stack.index(join.stack, tag="pred")$data
  variable.pred <- output$summary.fitted.values[index.pred,]
  colnames(variable.pred) <- sprintf("%s.%s", variable, colnames(variable.pred))
  sf.grid[colnames(variable.pred)] <- variable.pred

  # transform posterior marginal of precision into variance and calculate statistics
  i_field <- inla.spde2.result(output, name="spatial.field", spde)
  var.marginal <- data.frame(inla.zmarginal(inla.tmarginal(function(x) log(1/x), output$marginals.hyperpar[[1]]), silent=T)) %>%
    rename(mean=mean, sd=sd, `0.025quant`=quant0.025, `0.25quant`=quant0.25, `0.5quant`=quant0.5, `0.75quant`=quant0.75, `0.975quant`=quant0.975)
  # predictions
  cols <- sprintf(c("%s.mean", "%s.sd", "%s.0.025quant", "%s.0.5quant", "%s.0.975quant"), variable)
  i.pred <- sf.grid %>% as.data.frame %>% select(all_of(cols))
  colnames(i.pred) <- c("mean", "sd", "0.025quant", "0.5quant", "0.975quant")

  # Abiotic factor regression terms and predictions
  terms.abiotics[[variable]] <- rbind(
  	# Intercept
  	output$summary.fixed %>% 
  		mutate(ID=row.names(output$summary.fixed)) %>% remove_rownames %>% mutate(term="fixed") %>% 
  		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    # Precision
  	var.marginal %>% 
  		mutate(ID="log.variance.eps") %>% remove_rownames %>% mutate(term="spatial") %>% 
  		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
  	# Spatial component
  	bind_rows(list(log.variance=i_field$summary.log.variance.nominal, 
  	               log.range=i_field$summary.log.range.nominal), .id="ID") %>% 
  	  remove_rownames %>% mutate(term="spatial") %>% 
  		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    # Predictions
    cbind(sf.grid %>% as.data.frame %>% mutate(term="Predict", ID=Plot), i.pred) %>% 
    select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)
  	)
  i <- i + 1
}

terms.abiotics <- bind_rows(terms.abiotics, .id="model") %>% 
  mutate(species="", meta="Certain") %>%
  select(species, meta, model, term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)

write.table(terms.abiotics, "results_certain/abiotics.csv", sep = ",", col.names = T, row.names=F, append=F)
```

Visualization of predicted abiotic variables

```{r}
for (var in c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")) {
  var.mean <- sprintf("%s.mean", var)
  var.sd <- sprintf("%s.sd", var)

  # Take predicted abiotic in the grid and observed in the plot
  results <- sf.grid %>% select(geometry.1, var.mean,  var.sd) %>% st_set_geometry("geometry.1") %>%
    st_join(sf.plots %>% select(geometry, var), st_contains) %>% group_by(geometry.1) %>% summarize(
      y.mean=mean(get(var.mean), na.rm=T), 
      y.sd=mean(get(var.sd), na.rm=T), 
      y=mean(get(var), na.rm=T)
      )

  # Plot (ggplot)
  plot1 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=y), data=results) + 
    scale_fill_gradient2(low='blue', midpoint=0, high='red', limits=c(-3,3)) + 
    ggtitle(sprintf('Observed %s', var)) + theme(legend.position="bottom")
  plot2 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=y.mean), data=results) + 
    scale_fill_gradient2(low='blue', midpoint=0, high='red', limits=c(-3,3)) + 
    ggtitle(sprintf('Predicted %s', var)) + theme(legend.position="bottom")
  plot3 <- ggplot(NL) + geom_sf() + geom_sf(aes(geometry=geometry.1, fill=y.sd), data=results) + 
    scale_fill_gradient(low='white', high='red', limits=c(0,1)) + #scale_fill_gradient(low='darkblue', high='yellow', limits=c(0,1)) + 
    ggtitle(sprintf('Uncertainty %s', var)) + theme(legend.position="bottom")
  plot_grid(plot1, plot2, plot3, nrow = 1)
}
```

Fit the two-stage model using predicted abiotic variables, save model parameters and predictions

```{r}

for (species.take in species) { #c("Gestreepte witbol")
  print(species.take)

  start <- Sys.time()
  stk_y <- inla.stack(data = list(y=sf.plots[[species.take]]),
                A = list(A_y, 1, 1, 1, 1, 1, 1, 1, 1), 
                effects = list(c(list(u.field=1:spde$n.spde), list(intercept=1)), 
                               list(terrain=sf.plots$terrain), 
                               list(Dagnummer.z=sf.plots$Dagnummer.z), 
                               list(pH.z=sf.plots$pH.z), 
                               list(ORG_STOF.z=sf.plots$ORG_STOF.z.mean),
                               list(CN.z=sf.plots$CN.z.mean), 
                               list(N_totaal.z=sf.plots$N_totaal.z.mean), 
                               list(P_totaal.z=sf.plots$P_totaal.z.mean), 
                               list(K_CaCl2.z=sf.plots$K_CaCl2.z.mean)
                               ),
                tag="est.y")

  n <- nrow(sf.grid)
  stk_pred <- inla.stack(data=list(y=rep(NA, n)),  
                          effects=list(c(list(u.field=1:spde$n.spde), list(intercept=1)),
                                       list(terrain=sf.grid$terrain), 
                                       list(Dagnummer.z=sf.grid$Dagnummer.z), 
                                       list(pH.z=sf.grid$pH.z), 
                                       list(ORG_STOF.z=sf.grid$ORG_STOF.z.mean),
                                       list(CN.z=sf.grid$CN.z.mean), 
                                       list(N_totaal.z=sf.grid$N_totaal.z.mean), 
                                       list(P_totaal.z=sf.grid$P_totaal.z.mean), 
                                       list(K_CaCl2.z=sf.grid$K_CaCl2.z.mean)
                               ),
                          A=list(A_pred, 1, 1, 1, 1, 1, 1, 1, 1),
                          tag="est.pred")

  stk <- inla.stack(stk_y, stk_pred)

  formula <- y ~ -1 + intercept + f(inla.group(terrain), model = "rw2") + 
    f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") +
    CN.z + ORG_STOF.z + N_totaal.z + P_totaal.z + K_CaCl2.z + f(u.field, model=spde)

  output <-  tryCatch(
        {inla(formula, num.threads = 1,
                  data=inla.stack.data(stk, spde=spde), family="binomial",
                  control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=1), #
                  control.inla = list(strategy = "adaptive", int.strategy = "eb")) #int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

  # Model results
  if (!is.null(output)) {
    u_field <- inla.spde2.result(output, name="u.field", spde)
    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    # SDM regression terms and predictions
    terms.sdm <- rbind(
      # Splines for day number, pH, terrain
    	bind_rows(output$summary.random, .id="term") %>% 
    			filter(term %in% c("inla.group(Dagnummer.z)", "inla.group(pH.z)", "inla.group(terrain)")) %>% 
    	    mutate(term=str_replace_all(term, c("inla.group\\("="", "\\)"=""))) %>% 
    			select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    	# Linear terms for CN, ORG_STOF, N_totaal, P_totaal, K_CaCl2
    	output$summary.fixed %>% 
    		mutate(ID=row.names(output$summary.fixed)) %>% remove_rownames %>% mutate(term="fixed") %>% 
    		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    	# Spatial component
    	bind_rows(list(log.variance=u_field$summary.log.variance.nominal, 
    	               log.range=u_field$summary.log.range.nominal), .id="ID") %>% 
    	  remove_rownames %>% mutate(term="spatial") %>% 
    		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
      # Predictions
    	cbind(sf.grid %>% as.data.frame %>% mutate(term="Predict", ID=Plot), 
                       output$summary.fitted.values[index.pred,]) %>% 
      select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
      # Predictions (log-odds scale)
    	cbind(sf.grid %>% as.data.frame %>% mutate(term="APredict", ID=Plot), 
                       output$summary.linear.predictor[index.pred,]) %>% 
      select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)
    ) %>% 
      mutate(species=species.take, meta="Certain", model = "SDM") %>%
      select(species, meta, model, term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)

    # Save
    terms <- rbind(terms.sdm)#, terms.abiotics %>% mutate(species=species.take)
    fn <- sprintf('results_certain/%s.csv', species.take)
    write.table(terms, fn, sep = ",", col.names = T, row.names=F, append=F)
    #write.table(terms, "results_certain.csv", sep = ",", col.names = !file.exists("results_certain.csv"), row.names=F, append=T)
  }

  end <- Sys.time()
  print(end-start)
}
```

### Joint model

Construct a separate model for each abiotic variable:

```{r}
variables <- c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")

i <- 1
stacks <- list()
for (var in variables) {
  # Observations of the variable
  sf.var <- sf.plots %>% filter(!is.na(get(var)))
  A_var <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.var %>% as.data.frame %>% select(X,Y)))
  # One output for every estimated variable (+ species occurrence)
  y <- matrix(nrow=nrow(sf.var), ncol=length(variables) + 1)
  y.0 <- matrix(nrow=nrow(sf.grid), ncol=length(variables) + 1)
  y[,i] <- sf.var[[var]]
  # Stack
  effects <- list()
  effects[[sprintf("%s.i.field", var)]] = 1:spde$n.spde
  effects[[sprintf("%s.intercept", var)]] = 1 
  stacks[[var]] <- inla.stack(data=list(y=y), effects=list(effects), A=list(A_var), tag=sprintf("est.%s", var))
  stacks[[sprintf("%s.pred", var)]] <- inla.stack(data=list(y=y.0), effects=list(effects), A=list(A_pred), tag=sprintf("est.pred.%s", var))
  i <- i + 1
}
```

Fit the joint model using models for abiotic variables, save model parameters and predictions

```{r}

for (species.take in species) { #c("Gestreepte witbol")
  print(species.take)

  start <- Sys.time()
  # Species occurrence
  y <- matrix(nrow=nrow(sf.plots), ncol=length(variables) + 1)
  y[,i] <- sf.plots[[species.take]]
  # Stack
  stk_y <- inla.stack(data=list(y=y),  
                          effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                              CN.z.field=1:spde$n.spde,
                                              N_totaal.z.field=1:spde$n.spde, 
                                              P_totaal.z.field=1:spde$n.spde, 
                                              K_CaCl2.z.field=1:spde$n.spde,
                                              u.field=1:spde$n.spde), list(intercept=1)),
                                       list(terrain=sf.plots$terrain), 
                                       list(Dagnummer.z=sf.plots$Dagnummer.z), 
                                       list(pH.z=sf.plots$pH.z)),
                          A=list(A_y, 1, 1, 1),
                          tag="est.y")
  # Predicted over grid
  y <- matrix(nrow=nrow(sf.grid), ncol=length(variables) + 1)
  stk_pred <- inla.stack(data=list(y=y),  
                          effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                              CN.z.field=1:spde$n.spde,
                                              N_totaal.z.field=1:spde$n.spde, 
                                              P_totaal.z.field=1:spde$n.spde, 
                                              K_CaCl2.z.field=1:spde$n.spde,
                                              u.field=1:spde$n.spde), list(intercept=1)),
                                       list(terrain=sf.grid$terrain), 
                                       list(Dagnummer.z=sf.grid$Dagnummer.z), 
                                       list(pH.z=sf.grid$pH.z)),
                          A=list(A_pred, 1, 1, 1),
                          tag="est.pred")

  stk <- inla.stack(stacks[["ORG_STOF.z"]], stacks[["ORG_STOF.z.pred"]],
                    stacks[["CN.z"]], stacks[["CN.z.pred"]], 
                    stacks[["N_totaal.z"]], stacks[["N_totaal.z.pred"]],
                    stacks[["P_totaal.z"]], stacks[["P_totaal.z.pred"]],
                    stacks[["K_CaCl2.z"]], stacks[["K_CaCl2.z.pred"]],
                    stk_y, stk_pred)

  formula <- y ~ -1 + intercept + ORG_STOF.z.intercept + CN.z.intercept + N_totaal.z.intercept + P_totaal.z.intercept + K_CaCl2.z.intercept +
    f(inla.group(terrain), model = "rw2") + f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") + 
    f(ORG_STOF.z.i.field, model=spde) + f(ORG_STOF.z.field, copy="ORG_STOF.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
    f(CN.z.i.field, model=spde) + f(CN.z.field, copy="CN.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
    f(N_totaal.z.i.field, model=spde) + f(N_totaal.z.field, copy="N_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
    f(P_totaal.z.i.field, model=spde) + f(P_totaal.z.field, copy="P_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
    f(K_CaCl2.z.i.field, model=spde) + f(K_CaCl2.z.field, copy="K_CaCl2.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
    f(u.field, model=spde)

  families <- c(rep("gaussian", length(variables)), "binomial")

  output <-  tryCatch(
        {inla(formula, family=families, num.threads=1,
                 data=inla.stack.data(stk), 
                 control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=i),
                 control.inla = list(strategy = "adaptive", int.strategy = "eb"))#int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

  # Model results
  if (!is.null(output)) {
    # random field and index in stack
    u_field <- inla.spde2.result(output, name="u.field", spde)
    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    # SDM regression terms and predictions
    terms.sdm <- rbind(
      # Splines for day number, pH, terrain
    	bind_rows(output$summary.random, .id="term") %>% 
    			filter(term %in% c("inla.group(Dagnummer.z)", "inla.group(pH.z)", "inla.group(terrain)")) %>% 
    	    mutate(term=str_replace_all(term, c("inla.group\\("="", "\\)"=""))) %>% 
    			select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    	# Intercept
    	output$summary.fixed %>% 
    		mutate(ID=row.names(output$summary.fixed)) %>% remove_rownames %>% mutate(term="fixed") %>% 
    		filter(ID  == "intercept") %>%
    		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    	# Linear terms for CN, ORG_STOF, N_totaal, P_totaal, K_CaCl2
    	output$summary.hyperpar %>% 
    		mutate(ID=row.names(output$summary.hyperpar)) %>% remove_rownames %>% mutate(term="fixed") %>% 
    		filter(ID %>% startsWith("Beta")) %>%
    	  mutate(ID=str_replace_all(ID, c("Beta for "="", ".field"=""))) %>%
    		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
    	# Spatial component
    	bind_rows(list(log.variance=u_field$summary.log.variance.nominal, 
    	               log.range=u_field$summary.log.range.nominal), .id="ID") %>% 
    	  remove_rownames %>% mutate(term="spatial") %>% 
    		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
      # Predictions
    	cbind(sf.grid %>% as.data.frame %>% mutate(term="Predict", ID=Plot), 
                       output$summary.fitted.values[index.pred,]) %>% 
      select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
      # Predictions (log-odds scale)
    	cbind(sf.grid %>% as.data.frame %>% mutate(term="APredict", ID=Plot), 
                       output$summary.linear.predictor[index.pred,]) %>% 
      select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)
    ) %>% 
      mutate(species=species.take, meta="Uncertain", model = "SDM") %>%
      select(species, meta, model, term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)

    j <- 1
    terms.abiotics <- list()
    for (var in variables) {
      ## transform posterior marginal of precision into variance and calculate statistics
      i_field <- inla.spde2.result(output, name=sprintf("%s.i.field", var), spde)
      var.marginal <- data.frame(inla.zmarginal(inla.tmarginal(function(x) log(1/x), output$marginals.hyperpar[[j]]), silent=T)) %>%
        rename(mean=mean, sd=sd, `0.025quant`=quant0.025, `0.25quant`=quant0.25, `0.5quant`=quant0.5, `0.75quant`=quant0.75, `0.975quant`=quant0.975)
      # predictions
      index.pred <- inla.stack.index(stk, tag=sprintf("est.pred.%s", var))$data
      i.pred <- output$summary.linear.predictor[index.pred,] %>% select(mean, sd, `0.025quant`, `0.5quant`,`0.975quant`)

      # Abiotic factor regression terms and predictions
      terms.abiotics[[var]] <- rbind(
      	# Intercept
      	output$summary.fixed %>% 
      		mutate(ID=row.names(output$summary.fixed)) %>% remove_rownames %>% mutate(term="fixed") %>% 
      		filter(ID  == sprintf("%s.intercept", var)) %>% mutate(ID = "intercept") %>% 
      		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
        # Precision
      	var.marginal %>% 
      		mutate(ID="log.variance.eps") %>% remove_rownames %>% mutate(term="spatial") %>% 
      		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
      	# Spatial component
      	bind_rows(list(log.variance=i_field$summary.log.variance.nominal, 
      	               log.range=i_field$summary.log.range.nominal), .id="ID") %>% 
      	  remove_rownames %>% mutate(term="spatial") %>% 
      		select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`),
        # Predictions
        cbind(sf.grid %>% as.data.frame %>% mutate(term="Predict", ID=Plot), i.pred) %>% 
        select(term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)
      )	
      j <- j + 1
    }
    terms.abiotics <- bind_rows(terms.abiotics, .id="model") %>% 
      mutate(species=species.take, meta="Uncertain") %>%
      select(species, meta, model, term, ID, mean, sd, `0.025quant`, `0.5quant`, `0.975quant`)

    # Save
    terms <- rbind(terms.sdm, terms.abiotics)
    fn <- sprintf('results_uncertain/%s.csv', species.take)
    write.table(terms, fn, sep = ",", col.names = T, row.names=F, append=F)
    #write.table(terms, "results_uncertain.csv", sep = ",", col.names = !file.exists("results_uncertain.csv"), row.names=F, append=T)

  }

  end <- Sys.time()
  print(end-start)
}
```

## Experiments: training and validation

### Provinces

Two-stage model (certain abiotics), save predictions and prevalences

```{r}
provincies <- sort(unique(sf.plots$naam))
for (species.take in species) { 
  print(species.take)
  for (provincie in provincies) {#c("Utrecht")
    print(provincie)
    sf.train = sf.plots %>% filter(naam != provincie) 
    sf.test = sf.plots %>% filter(naam == provincie) 
    A_train <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.train %>% as.data.frame %>% select(X,Y)))
    A_test <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.test %>% as.data.frame %>% select(X,Y)))

    # Fit
    stk_y <- inla.stack(data = list(y=sf.train[[species.take]]),
                  A = list(A_train, 1, 1, 1, 1, 1, 1, 1, 1), 
                  effects = list(c(list(u.field=1:spde$n.spde), list(intercept=1)), 
                                 list(terrain=sf.train$terrain), 
                                 list(Dagnummer.z=sf.train$Dagnummer.z), 
                                 list(pH.z=sf.train$pH.z), 
                                 list(ORG_STOF.z=sf.train$ORG_STOF.z.mean),
                                 list(CN.z=sf.train$CN.z.mean), 
                                 list(N_totaal.z=sf.train$N_totaal.z.mean), 
                                 list(P_totaal.z=sf.train$P_totaal.z.mean), 
                                 list(K_CaCl2.z=sf.train$K_CaCl2.z.mean)
                                 ),
                  tag="est.y")

    stk_pred <- inla.stack(data=list(y=rep(NA, nrow(sf.test))),  
                            effects=list(c(list(u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.test$terrain), 
                                         list(Dagnummer.z=sf.test$Dagnummer.z), 
                                         list(pH.z=sf.test$pH.z), 
                                         list(ORG_STOF.z=sf.test$ORG_STOF.z.mean),
                                         list(CN.z=sf.test$CN.z.mean), 
                                         list(N_totaal.z=sf.test$N_totaal.z.mean), 
                                         list(P_totaal.z=sf.test$P_totaal.z.mean), 
                                         list(K_CaCl2.z=sf.test$K_CaCl2.z.mean)
                                 ),
                            A=list(A_test, 1, 1, 1, 1, 1, 1, 1, 1),
                            tag="est.pred")

    stk <- inla.stack(stk_y, stk_pred)

    formula <- y ~ -1 + intercept + f(inla.group(terrain), model = "rw2") + 
      f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") +
      CN.z + ORG_STOF.z + N_totaal.z + P_totaal.z + K_CaCl2.z + f(u.field, model=spde)

    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    output <-  tryCatch(
        {inla(formula, num.threads=1,
                    data=inla.stack.data(stk, spde=spde), family="binomial",
                    control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=1), 
                    control.inla = list(strategy = "adaptive", int.strategy = "eb"), control.compute=list(config = TRUE)) #int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

    # First model results
    if (!is.null(output)) {
      # get inla posterior sample
      s1 <- inla.posterior.sample(400, output) 
      linear_predictor <- inla.posterior.sample.eval("APredictor", s1)[index.pred,]
      P = 1/(1+exp(-linear_predictor))
      sample_mean <- function(p) mean(rbinom(length(p), 1, p))
      area_prevalences <- apply(P, 2, sample_mean) #colMeans(1/(1+exp(-linear_predictor)))

      # Predicted vs observed
      predicted <- output$summary.fitted.values[index.pred,c("mean", "sd", "0.025quant", "0.975quant")]
      observed <- sf.test[[species.take]]

      predictions <- cbind(predicted, observed) %>% mutate(meta = "Certain", model="SDM", species=species.take, naam = provincie)
      prevalences <- data.frame(area_prevalences, meta = "Certain", model="SDM", species=species.take, naam = provincie)
      #write.table(predictions, "predictions_certain.csv", sep = ",", col.names = !file.exists("predictions_certain.csv"), row.names=F, append = T)
      #write.table(prevalences, "prevalences_certain.csv", sep = ",", col.names = !file.exists("prevalences_certain.csv"), row.names=F, append = T)
      fn.1 <- sprintf('predictions_certain/%s_%s.csv', species.take, provincie)
      fn.2 <- sprintf('prevalences_certain/%s_%s.csv', species.take, provincie)
      write.table(predictions, fn.1, sep = ",", col.names = T, row.names=F, append=F)
      write.table(prevalences, fn.2, sep = ",", col.names = T, row.names=F, append=F)
    }

  }
}
```

Joint model (uncertain abiotics), save predictions and prevalences

```{r}
provincies <- sort(unique(sf.plots$naam))
for (species.take in species) { 
  print(species.take)
  for (provincie in provincies) {#c("Utrecht")
    print(provincie)
    sf.train = sf.plots %>% filter(naam != provincie) 
    sf.test = sf.plots %>% filter(naam == provincie) 
    A_train <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.train %>% as.data.frame %>% select(X,Y)))
    A_test <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.test %>% as.data.frame %>% select(X,Y)))

    # Species occurrence
    y <- matrix(nrow=nrow(sf.train), ncol=length(variables) + 1)
    y[,i] <- sf.train[[species.take]]
    # Stack
    stk_y <- inla.stack(data=list(y=y),  
                            effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                                CN.z.field=1:spde$n.spde,
                                                N_totaal.z.field=1:spde$n.spde, 
                                                P_totaal.z.field=1:spde$n.spde, 
                                                K_CaCl2.z.field=1:spde$n.spde,
                                                u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.train$terrain), 
                                         list(Dagnummer.z=sf.train$Dagnummer.z), 
                                         list(pH.z=sf.train$pH.z)),
                            A=list(A_train, 1, 1, 1),
                            tag="est.y")
    # Predicted over grid
    y <- matrix(nrow=nrow(sf.test), ncol=length(variables) + 1)
    stk_pred <- inla.stack(data=list(y=y),  
                            effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                                CN.z.field=1:spde$n.spde,
                                                N_totaal.z.field=1:spde$n.spde, 
                                                P_totaal.z.field=1:spde$n.spde, 
                                                K_CaCl2.z.field=1:spde$n.spde,
                                                u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.test$terrain), 
                                         list(Dagnummer.z=sf.test$Dagnummer.z), 
                                         list(pH.z=sf.test$pH.z)),
                            A=list(A_test, 1, 1, 1),
                            tag="est.pred")

    stk <- inla.stack(stacks[["ORG_STOF.z"]], stacks[["CN.z"]], stacks[["N_totaal.z"]],
                      stacks[["P_totaal.z"]], stacks[["K_CaCl2.z"]], stk_y, stk_pred)

    formula <- y ~ -1 + intercept + ORG_STOF.z.intercept + CN.z.intercept + N_totaal.z.intercept + P_totaal.z.intercept + K_CaCl2.z.intercept +
      f(inla.group(terrain), model = "rw2") + f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") + 
      f(ORG_STOF.z.i.field, model=spde) + f(ORG_STOF.z.field, copy="ORG_STOF.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(CN.z.i.field, model=spde) + f(CN.z.field, copy="CN.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(N_totaal.z.i.field, model=spde) + f(N_totaal.z.field, copy="N_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
      f(P_totaal.z.i.field, model=spde) + f(P_totaal.z.field, copy="P_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
      f(K_CaCl2.z.i.field, model=spde) + f(K_CaCl2.z.field, copy="K_CaCl2.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(u.field, model=spde)

    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    families <- c(rep("gaussian", length(variables)), "binomial")
    output <-  tryCatch(
        {inla(formula, family=families, num.threads=1, #verbose=T,
              data=inla.stack.data(stk), 
              control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=i),
              control.inla = list(strategy = "adaptive", int.strategy = "eb"), control.compute=list(config = TRUE)) #int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

    # Second model results
    if (!is.null(output)) {
      # get inla posterior sample
      s1 <- inla.posterior.sample(400, output) 
      linear_predictor <- inla.posterior.sample.eval("APredictor", s1)[index.pred,]
      P = 1/(1+exp(-linear_predictor))
      sample_mean <- function(p) mean(rbinom(length(p), 1, p))
      area_prevalences <- apply(P, 2, sample_mean) #colMeans(1/(1+exp(-linear_predictor)))

      # Predicted vs observed
      predicted <- output$summary.fitted.values[index.pred,c("mean", "sd", "0.025quant", "0.975quant")]
      observed <- sf.test[[species.take]]

      predictions <- cbind(predicted, observed) %>% mutate(meta = "Uncertain", model="SDM", species=species.take, naam = provincie)
      prevalences <- data.frame(area_prevalences, meta = "Uncertain", model="SDM", species=species.take, naam = provincie)
      #write.table(predictions, "predictions_uncertain.csv", sep = ",", col.names = !file.exists("predictions_uncertain.csv"), row.names=F, append = T)
      #write.table(prevalences, "prevalences_uncertain.csv", sep = ",", col.names = !file.exists("prevalences_uncertain.csv"), row.names=F, append = T)
      fn.1 <- sprintf('predictions_uncertain/%s_%s.csv', species.take, provincie)
      fn.2 <- sprintf('prevalences_uncertain/%s_%s.csv', species.take, provincie)
      write.table(predictions, fn.1, sep = ",", col.names = T, row.names=F, append=F)
      write.table(prevalences, fn.2, sep = ",", col.names = T, row.names=F, append=F)
    }

  }
}
```

### FGR regions

Two-stage model (certain abiotics), save predictions and prevalences

```{r}
fgrs <- unique(sf.plots$Fys_geo_regio)
fgrs <- sort(fgrs[!(fgrs %in% c("gtw", "gtz"))]) # gtw (1) and gtz (5) samples
for (species.take in species) { 
  print(species.take)
  for (fgr in fgrs) {
    print(fgr)
    sf.train = sf.plots %>% filter(Fys_geo_regio != fgr) 
    sf.test = sf.plots %>% filter(Fys_geo_regio == fgr) 
    A_train <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.train %>% as.data.frame %>% select(X,Y)))
    A_test <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.test %>% as.data.frame %>% select(X,Y)))

    # Fit
    stk_y <- inla.stack(data = list(y=sf.train[[species.take]]),
                  A = list(A_train, 1, 1, 1, 1, 1, 1, 1, 1), 
                  effects = list(c(list(u.field=1:spde$n.spde), list(intercept=1)), 
                                 list(terrain=sf.train$terrain), 
                                 list(Dagnummer.z=sf.train$Dagnummer.z), 
                                 list(pH.z=sf.train$pH.z), 
                                 list(ORG_STOF.z=sf.train$ORG_STOF.z.mean),
                                 list(CN.z=sf.train$CN.z.mean), 
                                 list(N_totaal.z=sf.train$N_totaal.z.mean), 
                                 list(P_totaal.z=sf.train$P_totaal.z.mean), 
                                 list(K_CaCl2.z=sf.train$K_CaCl2.z.mean)
                                 ),
                  tag="est.y")

    stk_pred <- inla.stack(data=list(y=rep(NA, nrow(sf.test))),  
                            effects=list(c(list(u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.test$terrain), 
                                         list(Dagnummer.z=sf.test$Dagnummer.z), 
                                         list(pH.z=sf.test$pH.z), 
                                         list(ORG_STOF.z=sf.test$ORG_STOF.z.mean),
                                         list(CN.z=sf.test$CN.z.mean), 
                                         list(N_totaal.z=sf.test$N_totaal.z.mean), 
                                         list(P_totaal.z=sf.test$P_totaal.z.mean), 
                                         list(K_CaCl2.z=sf.test$K_CaCl2.z.mean)
                                 ),
                            A=list(A_test, 1, 1, 1, 1, 1, 1, 1, 1),
                            tag="est.pred")

    stk <- inla.stack(stk_y, stk_pred)

    formula <- y ~ -1 + intercept + f(inla.group(terrain), model = "rw2") + 
      f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") +
      CN.z + ORG_STOF.z + N_totaal.z + P_totaal.z + K_CaCl2.z + f(u.field, model=spde)

    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    output <-  tryCatch(
        {inla(formula, num.threads=1,
                    data=inla.stack.data(stk, spde=spde), family="binomial",
                    control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=1), 
                    control.inla = list(strategy = "adaptive", int.strategy = "eb"), control.compute=list(config = TRUE)) #int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

    # First model results
    if (!is.null(output)) {
      # get inla posterior sample
      s1 <- inla.posterior.sample(400, output) 
      linear_predictor <- inla.posterior.sample.eval("APredictor", s1)[index.pred,]
      P = 1/(1+exp(-linear_predictor))
      sample_mean <- function(p) mean(rbinom(length(p), 1, p))
      area_prevalences <- apply(P, 2, sample_mean) #colMeans(1/(1+exp(-linear_predictor)))

      # Predicted vs observed
      predicted <- output$summary.fitted.values[index.pred,c("mean", "sd", "0.025quant", "0.975quant")]
      observed <- sf.test[[species.take]]

      predictions <- cbind(predicted, observed) %>% mutate(meta = "Certain", model="SDM", species=species.take, naam = fgr)
      prevalences <- data.frame(area_prevalences, meta = "Certain", model="SDM", species=species.take, naam = fgr)
      #write.table(predictions, "predictions_certain2.csv", sep = ",", col.names = !file.exists("predictions_certain2.csv"), row.names=F, append = T)
      #write.table(prevalences, "prevalences_certain2.csv", sep = ",", col.names = !file.exists("prevalences_certain2.csv"), row.names=F, append = T)
      fn.1 <- sprintf('predictions_certain2/%s_%s.csv', species.take, fgr)
      fn.2 <- sprintf('prevalences_certain2/%s_%s.csv', species.take, fgr)
      write.table(predictions, fn.1, sep = ",", col.names = T, row.names=F, append=F)
      write.table(prevalences, fn.2, sep = ",", col.names = T, row.names=F, append=F)
    }

  }
}
```

Joint model (uncertain abiotics), save predictions and prevalences

```{r}
fgrs <- unique(sf.plots$Fys_geo_regio)
fgrs <- sort(fgrs[!(fgrs %in% c("gtw", "gtz"))]) # gtw (1) and gtz (5) samples
for (species.take in species) { 
  print(species.take)
  for (fgr in fgrs) {
    print(fgr)
    sf.train = sf.plots %>% filter(Fys_geo_regio != fgr) 
    sf.test = sf.plots %>% filter(Fys_geo_regio == fgr) 
    A_train <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.train %>% as.data.frame %>% select(X,Y)))
    A_test <- inla.spde.make.A(mesh=mesh, loc=as.matrix(sf.test %>% as.data.frame %>% select(X,Y)))

    # Species occurrence
    y <- matrix(nrow=nrow(sf.train), ncol=length(variables) + 1)
    y[,i] <- sf.train[[species.take]]
    # Stack
    stk_y <- inla.stack(data=list(y=y),  
                            effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                                CN.z.field=1:spde$n.spde,
                                                N_totaal.z.field=1:spde$n.spde, 
                                                P_totaal.z.field=1:spde$n.spde, 
                                                K_CaCl2.z.field=1:spde$n.spde,
                                                u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.train$terrain), 
                                         list(Dagnummer.z=sf.train$Dagnummer.z), 
                                         list(pH.z=sf.train$pH.z)),
                            A=list(A_train, 1, 1, 1),
                            tag="est.y")
    # Predicted over grid
    y <- matrix(nrow=nrow(sf.test), ncol=length(variables) + 1)
    stk_pred <- inla.stack(data=list(y=y),  
                            effects=list(c(list(ORG_STOF.z.field=1:spde$n.spde, 
                                                CN.z.field=1:spde$n.spde,
                                                N_totaal.z.field=1:spde$n.spde, 
                                                P_totaal.z.field=1:spde$n.spde, 
                                                K_CaCl2.z.field=1:spde$n.spde,
                                                u.field=1:spde$n.spde), list(intercept=1)),
                                         list(terrain=sf.test$terrain), 
                                         list(Dagnummer.z=sf.test$Dagnummer.z), 
                                         list(pH.z=sf.test$pH.z)),
                            A=list(A_test, 1, 1, 1),
                            tag="est.pred")

    stk <- inla.stack(stacks[["ORG_STOF.z"]], stacks[["CN.z"]], stacks[["N_totaal.z"]],
                      stacks[["P_totaal.z"]], stacks[["K_CaCl2.z"]], stk_y, stk_pred)

    formula <- y ~ -1 + intercept + ORG_STOF.z.intercept + CN.z.intercept + N_totaal.z.intercept + P_totaal.z.intercept + K_CaCl2.z.intercept +
      f(inla.group(terrain), model = "rw2") + f(inla.group(Dagnummer.z), model = "rw2") +  f(inla.group(pH.z), model = "rw2") + 
      f(ORG_STOF.z.i.field, model=spde) + f(ORG_STOF.z.field, copy="ORG_STOF.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(CN.z.i.field, model=spde) + f(CN.z.field, copy="CN.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(N_totaal.z.i.field, model=spde) + f(N_totaal.z.field, copy="N_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
      f(P_totaal.z.i.field, model=spde) + f(P_totaal.z.field, copy="P_totaal.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001))))  +
      f(K_CaCl2.z.i.field, model=spde) + f(K_CaCl2.z.field, copy="K_CaCl2.z.i.field", fixed=FALSE, hyper=list(theta=list(param=c(0, 0.001)))) +
      f(u.field, model=spde)

    index.pred <- inla.stack.index(stk, tag="est.pred")$data

    families <- c(rep("gaussian", length(variables)), "binomial")
    output <-  tryCatch(
        {inla(formula, family=families, num.threads=1, #verbose=T,
              data=inla.stack.data(stk), 
              control.predictor=list(A=inla.stack.A(stk), compute=TRUE, link=i),
              control.inla = list(strategy = "adaptive", int.strategy = "eb"), control.compute=list(config = TRUE)) #int.strategy = "eb"
        }, error = function(cond) 
            {message(cond)
            return(NULL)}
    )

    # Second model results
    if (!is.null(output)) {
      # get inla posterior sample
      s1 <- inla.posterior.sample(400, output) 
      linear_predictor <- inla.posterior.sample.eval("APredictor", s1)[index.pred,]
      P = 1/(1+exp(-linear_predictor))
      sample_mean <- function(p) mean(rbinom(length(p), 1, p))
      area_prevalences <- apply(P, 2, sample_mean) #colMeans(1/(1+exp(-linear_predictor)))

      # Predicted vs observed
      predicted <- output$summary.fitted.values[index.pred,c("mean", "sd", "0.025quant", "0.975quant")]
      observed <- sf.test[[species.take]]

      predictions <- cbind(predicted, observed) %>% mutate(meta = "Uncertain", model="SDM", species=species.take, naam = fgr)
      prevalences <- data.frame(area_prevalences, meta = "Uncertain", model="SDM", species=species.take, naam = fgr)
      #write.table(predictions, "predictions_uncertain2.csv", sep = ",", col.names = !file.exists("predictions_uncertain2.csv"), row.names=F, append = T)
      #write.table(prevalences, "prevalences_uncertain2.csv", sep = ",", col.names = !file.exists("prevalences_uncertain2.csv"), row.names=F, append = T)
      fn.1 <- sprintf('predictions_uncertain2/%s_%s.csv', species.take, fgr)
      fn.2 <- sprintf('prevalences_uncertain2/%s_%s.csv', species.take, fgr)
      write.table(predictions, fn.1, sep = ",", col.names = T, row.names=F, append=F)
      write.table(prevalences, fn.2, sep = ",", col.names = T, row.names=F, append=F)
    }

  }
}
```

## Combine and export data

Combine and export the results of fitting all models to data:

```{r}
# Model parameters and predictions in entire data set
c(list.files("results_certain", full.names=T),
  list.files("results_uncertain", full.names=T)) %>% 
  map_dfr(read_csv, show_col_types=F) %>% 
  mutate(species=replace_na(species, ""))  %>% 
  write.csv(file="data_paper/results.csv", row.names=F)
# Predicted occurrence and prevalence in validation data set (Provinces)
c(list.files("predictions_certain", full.names=T), 
  list.files("predictions_uncertain", full.names=T)) %>% 
  map_dfr(read_csv, show_col_types=F) %>% 
  mutate(species=replace_na(species, ""))  %>% 
  write.csv(file="data_paper/predictions_province.csv", row.names=F)
c(list.files("prevalences_certain", full.names=T), 
  list.files("prevalences_uncertain", full.names=T)) %>% 
  map_dfr(read_csv, show_col_types=F) %>% 
  mutate(species=replace_na(species, ""))  %>% 
  write.csv(file="data_paper/prevalences_province.csv", row.names=F)
# Predicted occurrence and prevalence in validation data set (FGR regions)
c(list.files("predictions_certain2", full.names=T), 
  list.files("predictions_uncertain2", full.names=T)) %>% 
  map_dfr(read_csv, show_col_types=F) %>% 
  mutate(species=replace_na(species, ""))  %>% 
  write.csv(file="data_paper/predictions_fgr.csv", row.names=F)
c(list.files("prevalences_certain2", full.names=T), 
  list.files("prevalences_uncertain2", full.names=T)) %>% 
  map_dfr(read_csv, show_col_types=F) %>% 
  mutate(species=replace_na(species, ""))  %>% 
  write.csv(file="data_paper/prevalences_fgr.csv", row.names=F)
```

## Results: crowberry example

```{r}
results <- read.csv("data_paper/results.csv")
#results <- read.csv("data_paper/results.csv", col.names=c("species", "meta", "model", "term", "ID", "mean", "sd", "X0.025quant", "X0.5quant", "X0.975quant")) # LSF cluster results
```

```{r}
species.take <- "Kraaihei" #Gewoon reukgras, Engels raaigras, Pitrus, Duizendbland, Jacobskruiskruid
terms <- results %>% filter((species == species.take) | (species == "")) #is.na(species)
```

Spline parameters (SDM)

```{r}
# Day number, pH, terrain
terms.plot <- terms %>% filter(term %in% c("pH.z", "Dagnummer.z", "terrain")) %>% mutate(ID=ID %>% as.numeric) %>%
  mutate(term=recode(term, "pH.z" = "pH (z-score)", "Dagnummer.z"="Day number (months)", "terrain"="Terrain (more - less natural)")) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint")))
plot.splines <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) + geom_line() +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + 
  scale_color_discrete(name = "Model") +
  scale_x_continuous(breaks=seq(-4,4), name='Covariate Value') + 
  scale_y_continuous(name='Value') +
  facet_wrap( ~ term, ncol = 10, scales="free")
plot.splines
```

Abiotic parameters (SDM)

```{r}
# Linear terms
terms.plot <- terms %>% filter((term == "fixed") & (ID != "intercept")) %>% 
  mutate(ID=fct_relevel(ID, "ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")) %>% #"intercept", 
  mutate(term=recode(term, "fixed" = "Abiotic (z-score)")) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint")))
plot.abiotic <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + 
  scale_color_discrete(name = "Model Abiotics") +
  scale_x_discrete(name='Covariate') + 
  scale_y_continuous(name='Value') + #coord_cartesian(ylim=c(-5,5))+ #ylim=c(-10,10)
  facet_wrap( ~ term, ncol = 10, scales="free")
plot.abiotic
```

Spatial component parameters (SDM)

```{r}
# Spatial terms
terms.plot <- terms %>% filter((model == "SDM") & (term == "spatial")) %>% 
  mutate(ID=fct_relevel(ID, "log.range", "log.variance")) %>%
  mutate(term=recode(term, "spatial" = "Spatial (km)")) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint")))
plot.spatial <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + 
  scale_color_discrete(name = "Model Abiotics") +
  scale_x_discrete(name='Covariate') + 
  scale_y_continuous(name='Value') + #+coord_cartesian(ylim=c(-5,5))+ #ylim=c(-10,10)
  facet_wrap( ~ term, ncol = 10, scales="free")
plot.spatial
```

Spatial component parameters (abiotics)

```{r}
# Spatial terms (abiotics)
terms.plot <- terms %>% filter((term =="spatial") & (model != "SDM")) %>%
  mutate(ID=fct_relevel(ID, "log.range", "log.variance", "log.variance.eps")) %>%
  mutate(term=recode(term, "spatial" = "Spatial (km)"))%>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint")))
plot.extrapolate <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + 
  scale_color_discrete(name = "Model Abiotics") +
  scale_x_discrete(name='Covariate') + 
  scale_y_continuous(name='Value') +
  facet_wrap( ~ model, ncol = 10, scales="free")
plot.extrapolate
```

Combine all parameters to a single plot

```{r}
plot.all <- ggdraw() +
  draw_plot(plot.splines, x = 0, y = .666, width = 1, height = .333) +
  draw_plot(plot.abiotic + theme(legend.position = "none"), x = 0, y = .333, width = .77, height = .333) +
  draw_plot(plot.spatial + theme(legend.position = "none"), x = 0.77, y = .333, width = 0.23, height = 0.333) +
  draw_plot(plot.extrapolate + theme(legend.position = "none"), x = 0, y = 0, width = 1, height = 0.333) +
  draw_plot_label(label = c("A", "B", "C", "D"), size = 15, x = c(0.03, 0.03, 0.79, 0.03), y = c(0.96, 0.70, 0.70, 0.37))
plot.all
ggsave('results_paper/species_parameters.png', width=2000, heigh=2500, units='px')
```

Map of observed vs. predicted SDM from both models, probability scale

```{r}
# Predictions
terms.plot <- terms %>% filter((model == "SDM") & (term == "Predict")) %>% mutate(Plot=ID %>% as.factor)

# Combine data observations with model predictions
sf.base <- sf.grid %>% select(Plot, geometry.1) %>% st_set_geometry("geometry.1")
terms.y <- sf.base %>% st_join(sf.plots %>% select(all_of(species.take)), st_contains) %>% 
  group_by(Plot) %>% summarize(y=mean(get(species.take), na.rm=T)) %>% as.data.frame %>% select(Plot, y)
terms.sf <- sf.base %>% merge(terms.plot) %>% merge(terms.y)

plot0 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=y), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_viridis_c(option="B", limits=c(0,1)) + ggtitle("Raw data") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot1 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_viridis_c(option="B", limits=c(0,1)) + ggtitle("Joint Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot2 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_viridis_c(option="B", limits=c(0,1)) + ggtitle("Two-Stage Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot4 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_gradient(low='darkblue', high='yellow', limits=c(0,max(terms.sf$sd))) + ggtitle("Joint Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot5 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_gradient(low='darkblue', high='yellow', limits=c(0,max(terms.sf$sd))) + ggtitle("Two-Stage Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
#plot.all <- plot_grid(plot0, plot1, plot2, NULL, plot4, plot5, nrow = 2) + theme(plot.background = element_rect(fill = "white", colour = NA))
plot.all <- ggdraw() +
  draw_plot(plot0, x = 0, y = .5, width = 0.33, height = .5) +
  draw_plot(plot1, x = 0.33, y = .5, width = 0.33, height = .5) +
  draw_plot(plot2, x = 0.66, y = .5, width = 0.33, height = .5) +
  draw_plot(plot4, x = 0.33, y = 0, width = 0.33, height = .5) +
  draw_plot(plot5, x = 0.66, y = 0, width = 0.33, height = .5) +
  draw_plot_label(label = c("A", "B", "C", "D", "E"), size = 15, x = c(0.00, 0.33, 0.66, 0.33, 0.66), y = c(1.00, 1.00, 1.00, 0.50, 0.50))
plot.all
ggsave('results_paper/species_SDM_y.png', width=3200, heigh=3000, units='px')
```

Map of observed vs. predicted SDM from both models, log-odds scale

```{r}
# Predictions (log-odds)
terms.plot <- terms %>% filter((model == "SDM") & (term == "APredict")) %>% mutate(Plot=ID %>% as.factor)
y.mean <- mean(sf.plots[[species.take]], na.rm=T)
y.mean <- log(y.mean/(1-y.mean))

# Combine data observations with model predictions
sf.base <- sf.grid %>% select(Plot, geometry.1) %>% st_set_geometry("geometry.1")
terms.y <- sf.base %>% st_join(sf.plots %>% select(all_of(species.take)), st_contains) %>% 
  group_by(Plot) %>% summarize(y=mean(get(species.take), na.rm=T)) %>% as.data.frame %>% select(Plot, y)
terms.sf <- sf.base %>% merge(terms.plot) %>% merge(terms.y)

plot0 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=y), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_viridis_c(option="B", limits=c(0,1)) + ggtitle("Raw data") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot1 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_continuous_divergingx(palette = 'PiYG', mid = y.mean) + ggtitle("Joint Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot2 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_continuous_divergingx(palette = 'PiYG', mid = y.mean) + ggtitle("Two-Stage Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot4 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_gradient(low='darkblue', high='orange', limits=c(0,max(terms.sf$sd))) + ggtitle("Joint Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot5 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_gradient(low='darkblue', high='orange', limits=c(0,max(terms.sf$sd))) + ggtitle("Two-Stage Model") + 
  theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
#plot.all <- plot_grid(plot0, plot1, plot2, NULL, plot4, plot5, nrow = 2) + theme(plot.background = element_rect(fill = "white", colour = NA))
plot.all <- ggdraw() +
  draw_plot(plot0, x = 0, y = .5, width = 0.33, height = .5) +
  draw_plot(plot1, x = 0.33, y = .5, width = 0.33, height = .5) +
  draw_plot(plot2, x = 0.66, y = .5, width = 0.33, height = .5) +
  draw_plot(plot4, x = 0.33, y = 0, width = 0.33, height = .5) +
  draw_plot(plot5, x = 0.66, y = 0, width = 0.33, height = .5) +
  draw_plot_label(label = c("A", "B", "C", "D", "E"), size = 15, x = c(0.00, 0.33, 0.66, 0.33, 0.66), y = c(1.00, 1.00, 1.00, 0.50, 0.50))
plot.all
ggsave('results_paper/species_SDM_logodds.png', width=3200, heigh=3000, units='px')

```

Map of observed vs. predicted abiotics from both models

```{r}
for (var in c("ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")) {#var <- "ORG_STOF.z"
# Predictions
terms.plot <- terms %>% filter((model == var) & (term == "Predict")) %>% mutate(Plot=ID %>% as.factor)

# Combine data observations with model predictions
sf.base <- sf.grid %>% select(Plot, geometry.1) %>% st_set_geometry("geometry.1")
terms.y <- sf.base %>% st_join(sf.plots[var], st_contains) %>% 
  group_by(Plot) %>% summarize(y=mean(get(var), na.rm=T)) %>% as.data.frame %>% select(Plot, y)
terms.sf <- sf.base %>% merge(terms.plot) %>% merge(terms.y)

plot0 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=y), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_gradient2(low='blue', midpoint=0, high='red', limits=c(-3,3)) + 
  ggtitle(sprintf('Raw data %s', var)) + theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot1 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_gradient2(low='blue', midpoint=0, high='red', limits=c(-3,3)) + 
  ggtitle(sprintf('Joint Model %s', var)) + theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot2 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=mean), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_gradient2(low='blue', midpoint=0, high='red', limits=c(-3,3)) + 
  ggtitle(sprintf('Two-Stage Model %s', var)) + theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot4 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Uncertain")) + 
  scale_fill_gradient(low='white', high='red', limits=c(0,max(terms.sf$sd))) + 
  ggtitle(sprintf('Joint Model %s', var)) + theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
plot5 <- ggplot(NL) + geom_sf() + geom_sf(aes(fill=sd), data=terms.sf %>% filter(meta=="Certain")) + 
  scale_fill_gradient(low='white', high='red', limits=c(0,max(terms.sf$sd))) + 
  ggtitle(sprintf('Two-Stage Model %s', var)) + theme(legend.position="bottom", legend.key.width = unit(0.06, "npc"))
#plot.all <- plot_grid(plot0, plot1, plot2, NULL, plot4, plot5, nrow = 2) + theme(plot.background = element_rect(fill = "white", colour = NA))
plot.all <- ggdraw() +
  draw_plot(plot0, x = 0, y = .5, width = 0.33, height = .5) +
  draw_plot(plot1, x = 0.33, y = .5, width = 0.33, height = .5) +
  draw_plot(plot2, x = 0.66, y = .5, width = 0.33, height = .5) +
  draw_plot(plot4, x = 0.33, y = 0, width = 0.33, height = .5) +
  draw_plot(plot5, x = 0.66, y = 0, width = 0.33, height = .5) +
  draw_plot_label(label = c("A", "B", "C", "D", "E"), size = 15, x = c(0.00, 0.33, 0.66, 0.33, 0.66), y = c(1.00, 1.00, 1.00, 0.50, 0.50))
ggsave(sprintf('results_paper/species_%s.png', var), width=3200, heigh=3000, units='px')
}
```

Scatter plot of predicted values in each grid cell: two-stage vs joint model

```{r}
terms.plot.1 <- terms %>% filter((model == "SDM") & (term == "APredict")) %>% mutate(Plot=ID %>% as.factor) %>% 
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage Model", "Uncertain"="Joint Model"), levels=c("Two-Stage Model", "Joint Model"))) %>%
  select(species,meta,model,term,ID,mean,sd) %>% pivot_longer(c(mean,sd), names_to="type", values_to="estimate") %>% 
  pivot_wider(names_from = meta, values_from = estimate)

terms.plot.2 <- terms %>% filter((model != "SDM") & (term == "Predict")) %>% 
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage Model", "Uncertain"="Joint Model"), levels=c("Two-Stage Model", "Joint Model"))) %>%
  select(meta,model,term,ID,mean,sd) %>% pivot_longer(c(mean,sd), names_to="type", values_to="estimate") %>% 
  pivot_wider(names_from = meta, values_from = estimate) %>% mutate(species=species.take)

terms.plot <- rbind(terms.plot.1, terms.plot.2) %>% 
  mutate(model=fct_relevel(model, "SDM", "ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z"))

plot.scatter <- ggplot() + geom_point(aes(x=`Two-Stage Model`, y=`Joint Model`), data=terms.plot, size=0.1, alpha=0.2) + geom_abline() + 
     facet_wrap( ~ type + model, ncol = 3, scales="free") #+ ggtitle('Predicted Log-Odds (SDM) and \nPredicted Mean (Abiotics) at Grid Cells')
plot.scatter
ggsave('results_paper/species_scatter.png', width=1800, heigh=2400, units='px')
```

## Results: all species

Abiotic parameters for all species two-stage vs. joint model

```{r}
# Linear terms for all species
terms.plot <- results %>% filter((term == "fixed") & (model == "SDM")  & (ID != "intercept")) %>% 
  mutate(ID=fct_relevel(ID, "ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z")) %>% #"intercept",  
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "Joint"))) %>%
  merge(species.map) %>% mutate(species = name)
plot.abiotics.all <- ggplot(terms.plot, aes(x=ID, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) +
  geom_errorbar(position=position_dodge(width=0.5), aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + 
  scale_color_discrete(name = "Model") +
  scale_x_discrete(name='Covariate') + 
  scale_y_continuous(name='Value') +coord_cartesian(ylim=c(-5,5))+ #ylim=c(-10,10)
  facet_wrap( ~ species, ncol = 5)#, scales="free"
plot.abiotics.all
ggsave('results_paper/all_abiotics.png', width=2400, heigh=4000, units='px')
```

Table of correlations between predictions in each grid cell of two-stage vs. joint model

```{r}
# Uncertain vs Certain abiotic predictions
certain.abiotics <- results %>% filter(species == "") %>% transmute(model,ID,Certain=mean)
uncertain.abiotics <- results %>% filter((meta == "Uncertain") & (model != "SDM") & (term == "Predict")) %>% transmute(species, model,ID,Uncertain=mean)
abiotics <- uncertain.abiotics %>% merge(certain.abiotics)
# Uncertain vs Certain SDM predictions
certain.sdm <- results %>% filter((meta == "Certain") & (model == "SDM") & (term == "APredict")) %>% transmute(species, model,ID,Certain=mean)
uncertain.sdm <- results %>% filter((meta == "Uncertain") & (model == "SDM") & (term == "APredict")) %>% transmute(species, model,ID,Uncertain=mean)
sdm <- certain.sdm %>% merge(uncertain.sdm)
# Table of correlations
compare <- rbind(abiotics, sdm)
compare.cor <- compare %>% group_by(species, model) %>% 
  summarize(cor = sprintf("%.2f", round(cor(Certain, Uncertain),2))) %>% 
  pivot_wider(names_from=model, values_from=cor) %>%
  merge(species.map) %>% mutate(species = name) %>% select(-name) %>% arrange(species)
compare.cor %>% write.table('results_paper/correlation.csv', sep=',', row.names=F)
```

Histogram of these correlations

```{r}
compare.cor <- compare %>% group_by(species, model) %>% 
  summarize(cor = cor(Certain, Uncertain)) %>%
  mutate(model=fct_relevel(model, "SDM", "ORG_STOF.z", "CN.z", "N_totaal.z", "P_totaal.z", "K_CaCl2.z"))

ggplot(compare.cor, aes(x=cor)) + 
  geom_histogram(color="black", fill="white", breaks=seq(0.8,1.0,0.01)) + 
  ylab("Number of Species") +xlab("Pearson Correlation Coefficient") +
  facet_wrap( ~ model, ncol = 3)  #+ggtitle('Correlation between Certain vs. Uncertain Abiotic Model Predictions')
ggsave('results_paper/species_correlation.png', width=2000, heigh=1200, units='px')

```

Table of correlations between abiotic parameters in each species of two-stage vs. joint model

```{r}
# Uncertain vs Certain abiotic predictions
certain.abiotics <- results %>% filter((meta == "Certain") & (model == "SDM") & (term == "fixed") & (ID != "intercept")) %>%
  transmute(species, model,ID,Certain=mean)
uncertain.abiotics <- results %>% filter((meta == "Uncertain") & (model == "SDM") & (term == "fixed") & (ID != "intercept")) %>%
  transmute(species, model,ID,Uncertain=mean)
abiotics <- uncertain.abiotics %>% merge(certain.abiotics)
# Table of correlations
compare.cor <- abiotics %>% group_by(ID) %>% summarize(cor = round(cor(Certain, Uncertain),2)) %>% 
  pivot_wider(names_from=ID, values_from=cor)
compare.cor
```

Validation set illustration

```{r}
# Leave province out
sf.grid$provincie <- as.factor(sf.grid$naam)
plot.validation.1 <- ggplot(sf.grid) +
	geom_sf(aes(geometry=geometry.1, fill=provincie)) + 
	geom_sf(fill=NA, data=NL) +
  scale_fill_discrete(name = "Province") +
	ggtitle('Leave-province-out')

# Leave FGR region out
sf.grid$fgr <- factor(sf.grid$lmf_indeling, levels=c("afz", "duo", "duw", "hll", "hzn", "hzo", "hzv",
                                                     "hzz", "lvh", "lvn", "riv", "zkm", "zkn", "zkz"))
plot.validation.2 <- ggplot(sf.grid) +
	geom_sf(aes(geometry=geometry.1, fill=fgr)) + 
	geom_sf(fill=NA, data=NL) +
  scale_fill_manual(name = "FGR", values=c("afz"="lightseagreen", "duo"="darkgoldenrod", "duw"="darkgoldenrod1", "hll"="forestgreen", 
                                           "hzn"="wheat1", "hzo"="wheat2", "hzv"="wheat3", "hzz"="wheat4", 
                                           "lvh"="darkseagreen", "lvn"="darkseagreen1", "riv"="steelblue", 
                                           "zkm"="snow2", "zkn"="snow3", "zkz"="snow4")) +
	ggtitle('Leave-FGR-out')

# Combine plots
plot.all <- ggdraw() +
  draw_plot(plot.validation.1, x = 0, y = 0, width = 0.53, height = 1) +
  draw_plot(plot.validation.2, x = 0.53, y = 0, width = 0.47, height = 1) +
  draw_plot_label(label = c("A", "B"), size = 15, x = c(0.01, 0.53), y = c(0.99,0.99)) +
  theme(plot.background = element_rect(fill = "white", colour = NA))
plot.all
ggsave('results_paper/leave_out_validation.png', width=2500, heigh=1300, units='px')

```

## Results: training and validation

Calculate overall RMSE in each province

```{r}
predictions <- read.csv("data_paper/predictions_province.csv")
#predictions <- read.csv("data_paper/predictions_province.csv", col.names=c("mean", "sd", "X0.025quant", "X0.975quant", "observed", "meta", "model", "species", "naam")) # LSF cluster results
rmse.1 <- predictions %>% group_by(meta, species) %>% summarize(rmse=sqrt(mean((observed-mean)**2, na.rm=T))) 
```

Calculate overall RMSE in each FGR region

```{r}
predictions <- read.csv("data_paper/predictions_fgr.csv")
#predictions <- read.csv("data_paper/predictions_fgr.csv", col.names=c("mean", "sd", "X0.025quant", "X0.975quant", "observed", "meta", "model", "species", "naam")) # LSF cluster results
rmse.2 <- predictions %>% group_by(meta, species) %>% summarize(rmse=sqrt(mean((observed-mean)**2, na.rm=T))) 
```

Scatter plot of two-stage vs. joint model overall RMSE in each species:

```{r}
rmse.both <- rbind(rmse.1 %>% mutate(validation="Province"),
                   rmse.2 %>% mutate(validation="FGR")) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage Model", "Uncertain"="Joint Model"), levels=c("Two-Stage Model", "Joint Model")))
rmse.both %>% pivot_wider(names_from = c(validation, meta), values_from = rmse) %>% 
  mutate(across(2:5, function(x) sprintf("%.3f", round(x,3)))) %>% 
  merge(species.map) %>% mutate(species = name) %>% select(-name) %>% arrange(species)%>%
  write.table('results_paper/rmse_compare.csv', sep=',', row.names=F)

ggplot(rmse.both  %>% pivot_wider(names_from = meta, values_from = rmse)) + 
  geom_point(aes(x=`Two-Stage Model`, y=`Joint Model`), size=0.5) + geom_abline() +
  scale_x_continuous(trans='log10', labels = label_comma()) + scale_y_continuous(trans='log10', labels = label_comma()) +
   coord_cartesian(ylim=c(0.03,0.5), xlim=c(0.03,0.5)) + facet_wrap( ~ validation, ncol = 2) #+
  #ggtitle('Species validation RMSE: Certain vs. Uncertain Abiotics')
ggsave('results_paper/rmse_comparison.png', width=1200, heigh=800, units='px')

```

Compare predicted and observed prevalence in each province:

```{r}
prevalences <- read.csv("data_paper/prevalences_province.csv")
# prevalences <- read.csv("data_paper/prevalences_province.csv", col.names=c("area_prevalences", "meta", "model", "species", "naam")) # LSF cluster results

prevalences.predicted <- prevalences %>% group_by(meta, species, naam) %>% 
  summarize(mean = mean(area_prevalences, na.rm=T), 
            X0.025quant = quantile(area_prevalences, 0.025, na.rm=T),
            X0.975quant = quantile(area_prevalences, 0.975, na.rm=T))

prevalences.observed <- sf.plots %>% as.data.frame  %>% select(naam, all_of(species)) %>% 
  #mutate(naam=Fys_geo_regio) %>% filter(!(naam %in% c("gtw", "gtz"))) %>% 
  group_by(naam) %>% summarize_all(function(x) mean(x, na.rm=T)) %>% 
  pivot_longer(species, names_to = "species", values_to="mean") %>% mutate(meta="True")

prevalences.both <- rbind(prevalences.predicted, prevalences.observed) %>%
  filter((species != "Struikhei") & (!is.na(naam))) %>%
  mutate(naam=recode(naam, "Drenthe"='DR', "Flevoland"='FL', "Fryslân"='FR', "Gelderland"='GE', "Groningen"='GR',
                     "Limburg"='LI', "Noord-Brabant"='NB', "Noord-Holland"='NH', "Overijssel"='OV', "Utrecht"='UT',
                     "Zeeland"='ZE', "Zuid-Holland"='ZH')) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "True", "Joint"))) %>%
  merge(species.map) %>% mutate(species = name)

plot.all <- ggplot(prevalences.both, aes(x=naam, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) + 
  geom_errorbar(position=position_dodge(width=0.5), width=0.5, aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + #+ position_dodge()
  scale_color_discrete(name = "Model") +
  scale_x_discrete(name='Province') + 
  scale_y_continuous(name='Prevalence') +
  facet_wrap( ~ species, ncol = 5, scales='free_y')
plot.all
ggsave('results_paper/all_prevalences.png', width=2800, heigh=4000, units='px')
```

Compare predicted and observed prevalence in each FGR region:

```{r}
prevalences <- read.csv("data_paper/prevalences_fgr.csv")
# prevalences <- read.csv("data_paper/prevalences_fgr.csv", col.names=c("area_prevalences", "meta", "model", "species", "naam")) # LSF cluster results

prevalences.predicted <- prevalences %>% group_by(meta, species, naam) %>% 
  summarize(mean = mean(area_prevalences, na.rm=T), 
            X0.025quant = quantile(area_prevalences, 0.025, na.rm=T),
            X0.975quant = quantile(area_prevalences, 0.975, na.rm=T))

prevalences.observed <- sf.plots %>% as.data.frame  %>% select(Fys_geo_regio, all_of(species)) %>% 
  mutate(naam=Fys_geo_regio) %>% filter(!(naam %in% c("gtw", "gtz"))) %>% 
  group_by(naam) %>% summarize_all(function(x) mean(x, na.rm=T)) %>% 
  pivot_longer(species, names_to = "species", values_to="mean") %>% mutate(meta="True")

prevalences.both <- rbind(prevalences.predicted, prevalences.observed) %>%
  filter((species != "Struikhei") & (!is.na(naam))) %>%
  mutate(meta=factor(recode(meta, "Certain" = "Two-Stage", "Uncertain"="Joint"), levels=c("Two-Stage", "True", "Joint")))  %>%
  merge(species.map) %>% mutate(species = name) #%>%
  #mutate(naam=recode(naam, "Drenthe"='DR', "Flevoland"='FL', "Fryslân"='FR', "Gelderland"='GE', "Groningen"='GR',
  #                   "Limburg"='LI', "Noord-Brabant"='NB', "Noord-Holland"='NH', "Overijssel"='OV', "Utrecht"='UT',
  #                   "Zeeland"='ZE', "Zuid-Holland"='ZH'))

plot.all <- ggplot(prevalences.both, aes(x=naam, y=mean, col=meta, group=meta)) + 
  geom_point(position=position_dodge(width=0.5), size=2) + 
  geom_errorbar(position=position_dodge(width=0.5), width=0.5, aes(ymin=X0.025quant, ymax=X0.975quant)) +
  theme(legend.position="top", axis.text.x = element_text(angle = 90)) + #+ position_dodge()
  scale_color_discrete(name = "Model") +
  scale_x_discrete(name='Fys_geo_regio') + 
  scale_y_continuous(name='Prevalence') +
  facet_wrap( ~ species, ncol = 5, scales='free_y')
plot.all
ggsave('results_paper/all_prevalences2.png', width=2800, heigh=4000, units='px')
```