### 2023-08-15 paper_data.Rmd

---
title: "paper_data"
author: "Markus Viljanen"
date: "2023-08-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(readxl)
library(RPostgreSQL)
library(plyr)
library(tidyverse)
library(sf)
```

## Original LMF data set to LMF subset

Fetch Netherlands boundaries from RIVM GIS server

```{r}
conn <- dbConnect(PostgreSQL(), dbname="geodata", user = "ro", password="ro!geo", host = "geodatabase.rivm.nl")
NL <- st_read(conn, query = "SELECT * FROM nl.adm_wijkbuurt_2022_gen;") %>%
  filter(indeling == 'gemeente') %>% st_union()
st_write(NL, "data_paper/NL.shp", append=F)
```

Abiotic measurements for LMF data

```{r}
# Abiotic measurements for LMF data 
LMF.dir <- file.path('LMF', 'LMF data.xlsx')
df.abiotic <- read_excel(LMF.dir, 1)[-1,-1] %>% distinct(releve_nr, .keep_all=T) %>% 
  mutate(across(DATE:`N-(NO3+NO2)`,as.numeric)) %>% mutate(NAAM=tolower(str_replace_all(NAAM, " ", "")))
# Fix some identifier names in abiotic data
naam <- df.abiotic$NAAM
# some FR**** and FR**** identifiers were not properly zero padded
replace.index <- substr(naam,1,2) %in% c("fr", "fl")
pad.4 <- paste(substr(naam,1,2), str_pad(substr(naam,3,nchar(naam)), 4, pad="0"), sep="")
df.abiotic[replace.index, "NAAM"] <- pad.4[replace.index]
# Some GR**** identifiers were GR****/GR****
replace.index <- grepl('/', df.abiotic$NAAM, fixed = TRUE)
last.6 <- substr(naam, nchar(naam)-6+1, nchar(naam))
df.abiotic[replace.index, "NAAM"] <- last.6[replace.index]
# Lisa's fixes to coordinates 
# Tostrams, Lisa <lisa.tostrams@rivm.nl>
df.abiotic.2 <- read.csv('LMF/lmf_log_data_corrected_coords.csv') %>% 
  distinct(releve_nr, .keep_all=T) %>% 
  select(releve_nr, `X.coördinaat`, `Y.coördinaat`)
# Replace coordinates
df.abiotic <- df.abiotic %>% left_join(df.abiotic.2)  %>% 
  mutate(Plot = NAAM %>% factor, X = `X.coördinaat`, Y=`Y.coördinaat`) %>%
  group_by(Plot) %>% filter(DATE == max(DATE)) %>% filter(releve_nr != 203857) #NAAM = nb1019 has two releve_nr in same year
```

Original LMF data (Species names, Field visit information, Vegatation plots information, Species observations)

```{r}
# Original LMF data
# https://longreads.cbs.nl/meetprogrammas-flora-en-fauna-2022/

# Species
df.species = read.csv('LMF/LMF_Soorten.csv',sep=';', colClasses=c("numeric", "numeric", "factor", "factor")) %>% 
  select(Soortcode, Nednaam, Wetnaam)

# Visits
df.visit <- read.csv('LMF/LMF_kop.csv', sep=';', colClasses=c("factor", "numeric", "character", "numeric", "numeric", 
                                                              "factor", "factor", "factor", "factor", "factor", 
                                                              "factor", "factor", "factor", "factor", 
                                                              "numeric", "numeric", "numeric", "factor", "factor", "factor")) %>% 
  mutate(Datum=as.Date(Datum, format = "%Y%m%d")) %>% 
  select(Plot, Jaar, Dagnummer, Pq_grootte)

# Plots
df.plots <- read.csv('LMF/LMF_plot.csv',sep=';', dec=',', colClasses=c("factor", "numeric", "numeric", 
                                                                       "factor", "factor", "factor", "factor", "factor", 
                                                                       "factor", "factor", "factor", "factor", "numeric", "factor")) %>%
  mutate(X=X*1000, Y=Y*1000) %>%
  #mutate(x = as.numeric(gsub(",", ".", X))*1000,
  #       y = as.numeric(gsub(",", ".", Y))*1000) %>%
  select(Plot, X, Y, Fys_geo_regio)

# Observations, take only latest
df.occurrence <- read.csv('LMF/LMF_tellingen.csv', sep=';', colClasses=c("factor", "numeric", "numeric", "numeric")) %>% 
  select(Plot, Jaar, Soortcode, Bedekking)
df.occurrence <- df.occurrence %>% merge(df.species) %>% group_by(Plot) %>% filter(Jaar == max(Jaar))

# Add plot visit date 
df.extra <- df.occurrence %>% distinct(Plot, Jaar) %>% merge(df.visit)
df.plots <- df.plots %>% merge(df.extra)

# Add abiotic information
df.plots <- df.plots %>% full_join(df.abiotic, by="Plot", suffix = c("", ".abiotic")) %>%
  mutate(X=ifelse(is.na(X), X.abiotic, X),
         Y=ifelse(is.na(Y), Y.abiotic, Y))
```

Wieger's proposed species to analyze

```{r}
# Wamelink, Wieger <wieger.wamelink@wur.nl>
df.n <- df.occurrence %>% group_by(Nednaam) %>% count() %>% arrange(-n)
LMF.dir <- file.path('LMF', '8969892_WageningenEnvironmentalResearch_6708PB_20201006_ww3.xlsx')
species_proposal <- c(read_excel(LMF.dir)$NedNAAM, "Gestreepte witbol", "Struikhei")
species <- (data.frame(Nednaam=species_proposal) %>% merge(df.n) %>% filter(n>=20) %>% select(Nednaam))$Nednaam
chosen.species <- data.frame(Nednaam=species[species != "Struikhei"]) %>% 
  merge(df.species) %>% merge(df.n) %>% select(Nednaam, Wetnaam, n) %>% arrange(-n) 
```

Overlap statistics:

```{r}
c1 <- unique(df.abiotic$Plot)
c2 <- unique(df.visit$Plot)
cx <- intersect(c1, c2)
sprintf("Plots: %d abiotic, %d visit, %d intersection", length(c1), length(c2), length(cx))

c1 <- unique(df.abiotic$DATE)
c2 <- unique(df.visit$Jaar)
cx <- intersect(c1, c2)
sprintf("Years: %d abiotic, %d visit, %d intersection", length(c1), length(c2), length(cx))

c1 <- unique(paste(as.character(df.abiotic$Plot), 'x', as.character(df.abiotic$DATE)))
c2 <- unique(paste(as.character(df.visit$Plot), 'x', as.character(df.visit$Jaar)))
cx <- intersect(c1, c2)
sprintf("(Plot,Year)-pairs: %d abiotic, %d visit, %d intersection", length(c1), length(c2), length(cx))
```

Binary table of each species occurrence at each vegetation plot (0/1)

```{r}
df.y <- df.occurrence %>% select(Plot, Nednaam, Bedekking) %>% mutate(Bedekking=1) %>% spread(Nednaam, Bedekking, fill=0)
```

Take only relevant vegetation plot information and join to species occurrence at each plot

```{r}
df.plots <- df.plots %>% 
  select(Plot, releve_nr, X, Y, Fys_geo_regio, Jaar, Dagnummer, Pq_grootte, `pH-H20`, ORG_STOF, `C/N`, N_totaal, P_totaal, K_CaCl2) %>% 
  mutate(across(`pH-H20`:K_CaCl2, function (x) round(x, digits=2))) %>%
  full_join(df.y[c("Plot", species)])
```

## Add information, spatial joins with geodata

In codes we use X&Y as km to enforce a vague prior

```{r}
NL <- (NL - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
```

Data set of species observations and environmental conditions

```{r}
sf.plots <- df.plots %>% st_as_sf(coords = c("X","Y"), crs=st_crs(NL))
st_geometry(sf.plots) <- (sf.plots$geometry - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
sf.plots[c("X", "Y")] <- sf.plots  %>% st_coordinates()
```

Grid of all possible plots

```{r}
# Grid with rectangles and their centers  #%>% st_as_sf(crs=st_crs(NL))
grid.rect <- st_make_grid(NL,n=100,crs=st_crs(NL), what='polygons', square=T) 
grid.circle <- st_make_grid(NL,n=100,crs=st_crs(NL), what='centers')
index <- which(lengths(st_within(grid.circle, NL)) > 0) #st_intersects(grid.rect, NL)
# Create data frame
sf.grid <- data.frame(grid.circle[index], grid.rect[index]) %>% st_as_sf(crs=st_crs(NL))
sf.grid[c("X", "Y")] <- sf.grid %>% st_geometry %>% st_centroid() %>% st_coordinates()
sf.grid$Plot <- factor(sprintf("Grid%d", 1:nrow(sf.grid)))
```

Add GIS layers: province

```{r}
#https://service.pdok.nl/kadaster/bestuurlijkegebieden/atom/v1_0/bestuurlijke_gebieden.xml
provincie = st_read("BestuurlijkeGebieden_2023.gpkg", "provinciegebied") %>% select(naam)
st_geometry(provincie) <- (provincie$geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)

sf.plots <- sf.plots %>% st_join(provincie , left=T, st_within)
sf.grid <- sf.grid %>% st_join(provincie , left=T, st_within)
#plot(provincie)
```

Add GIS layers: FGR regio

```{r}
# File created by FGR.ipynb
fgr = st_read("FGR_lmf_mapping.gpkg") %>% select(lmf_indeling)
st_geometry(fgr) <- (fgr$geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)

sf.plots <- sf.plots %>% st_join(fgr , left=T, st_within)
sf.grid <- sf.grid %>% st_join(fgr , left=T, st_within)
#plot(fgr)
```

Add GIS layers: soiltype, landuse

```{r}
# https://www.nationaalgeoregister.nl/geonetwork/srv/dut/catalog.search#/metadata/2d3dd6d2-2d2b-4b5f-9e30-86e19ed77a56
# http://www.geodesk.nl/Grondsoorten.htm
## takes ~ 5 mins
landuse <- st_read("bestand_bodemgebruik_2015.gpkg", layer="bestand_bodemgebruik_2015")
soiltype <- st_read("Grondsoortenkaart-2006.shp")
landuse <- landuse %>% transmute(bodemgebruik=relevel(factor(bodemgebruik), ref="Bos"))
soiltype <- soiltype %>% transmute(OMSCHRIJVI=relevel(factor(OMSCHRIJVI), ref="Zand"))
st_geometry(landuse) <- (landuse$geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
st_geometry(soiltype) <- (soiltype$geometry - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
```

```{r}
## takes ~2 hours
info.plots <- sf.plots %>% select(Plot)
info.plots <- info.plots %>% st_join(landuse , left=T, st_within)
info.plots <- info.plots %>% st_join(soiltype , left=T, st_within)
info.plots <- info.plots %>% as.data.frame %>% distinct(Plot, bodemgebruik, OMSCHRIJVI)
#info.plots %>% saveRDS("plot.info.RDS")
#info.plots <- readRDS("plot.info.RDS")
sf.plots <- sf.plots %>% merge(info.plots, on="Plot")
```

```{r}
## takes ~2 hours
info.grid <- sf.grid %>% select(Plot)
info.grid <- info.grid %>% st_join(landuse , left=T, st_within)
info.grid <- info.grid %>% st_join(soiltype , left=T, st_within)
info.grid <- info.grid %>% as.data.frame %>% distinct(Plot, bodemgebruik, OMSCHRIJVI)
#info.grid %>% saveRDS("grid.info.RDS")
#info.grid <- readRDS("grid.info.RDS")
# sf.grid %>% merge(info.grid, on="Plot") drops the second geometry, so cbind the correctly sorted auxiliary data
sf.grid <- cbind(sf.grid, sf.grid %>% as.data.frame %>% select(Plot) %>% merge(info.grid, on="Plot", sort=F) %>% select(bodemgebruik, OMSCHRIJVI)) 
```

Fill very few NA observations with most common category and combine some categories

```{r}
sf.plots <- sf.plots %>% mutate(
  bodemgebruik= fct_na_value_to_level(fct_collapse(bodemgebruik, 
      Bebouwd=c("Bebouwd exclusief bedrijfsterrein", "Bedrijfsterrein", "Semi-bebouwd", "Glastuinbouw"),
      Spoorterrein=c("Spoorterrein", "Vliegveld")), "Landbouw en overig agrarisch"),
 OMSCHRIJVI=fct_na_value_to_level(relevel(factor(OMSCHRIJVI), ref="Zand"), "Zand"))

sf.grid <- sf.grid %>% mutate(
  bodemgebruik=fct_na_value_to_level(fct_collapse(bodemgebruik, 
      Bebouwd=c("Bebouwd exclusief bedrijfsterrein", "Bedrijfsterrein", "Semi-bebouwd", "Glastuinbouw"),
      Spoorterrein=c("Spoorterrein", "Vliegveld")), "Landbouw en overig agrarisch"),
  OMSCHRIJVI=fct_na_value_to_level(relevel(factor(OMSCHRIJVI), ref="Zand"), "Zand"))
```

Save data

```{r}
chosen.species %>% write.csv(file="data_paper/chosen_species.csv", row.names=F)
sf.plots %>% as.data.frame %>% select(-geometry) %>% write.csv(file="data_paper/data_plots.csv", row.names=F)
sf.grid %>% as.data.frame %>% select(-geometry,-geometry.1) %>% write.csv(file="data_paper/data_grid.csv", row.names=F)
```