### 2023-08-18 paper_data.Rmd
some review

---
title: "paper_data"
author: "Markus Viljanen"
date: "2023-08-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(readxl)
library(RPostgreSQL)
library(plyr)
library(tidyverse)
library(sf)
```

## Original LMF data set to LMF subset
Here as well as in the modelling code, all coordinates are in EPSG:28992 Projected coordinate system for Netherlands.

Abiotic measurements for LMF data

```{r}
# Abiotic measurements for LMF data 
LMF.dir <- file.path('LMF', 'LMF data.xlsx')
df.abiotic <- read_excel(LMF.dir, 1)[-1,-1] %>% distinct(releve_nr, .keep_all=T) %>% 
  mutate(across(DATE:`N-(NO3+NO2)`,as.numeric)) %>% mutate(NAAM=tolower(str_replace_all(NAAM, " ", "")))
# Fix some identifier names in abiotic data
naam <- df.abiotic$NAAM
# some FR**** and FR**** identifiers were not properly zero padded
replace.index <- substr(naam,1,2) %in% c("fr", "fl")
pad.4 <- paste(substr(naam,1,2), str_pad(substr(naam,3,nchar(naam)), 4, pad="0"), sep="")
df.abiotic[replace.index, "NAAM"] <- pad.4[replace.index]
# Some GR**** identifiers were GR****/GR****
replace.index <- grepl('/', df.abiotic$NAAM, fixed = TRUE)
last.6 <- substr(naam, nchar(naam)-6+1, nchar(naam))
df.abiotic[replace.index, "NAAM"] <- last.6[replace.index]
# Lisa's fixes to coordinates 
# This consisted of 24 plots of which the locations seemed to be outside of the Netherlands. The plotnames were found in the LMF plots, and for those the correct coordinates were taken from the LMF programme. 
# There were also an additional 329 abiotic plots with a plot name that was in the LMF file, but where the coordinates did not match. In those cases the coordinates from the LMF programme were taken as well. 
# Tostrams, Lisa <lisa.tostrams@rivm.nl>
df.abiotic.2 <- read.csv('LMF/lmf_log_data_corrected_coords.csv') %>% 
  distinct(releve_nr, .keep_all=T) %>% 
  select(releve_nr, `X.coördinaat`, `Y.coördinaat`)
# Replace coordinates
df.abiotic <- df.abiotic %>% left_join(df.abiotic.2)  %>% 
  mutate(Plot = NAAM %>% factor, X = `X.coördinaat`, Y=`Y.coördinaat`) %>%
  group_by(Plot) %>% filter(DATE == max(DATE)) %>% filter(releve_nr != 203857) #NAAM = nb1019 has two releve_nr in same year
```

Original LMF data (Species names, Field visit information, Vegatation plots information, Species observations)

```{r}
# Original LMF data
# https://longreads.cbs.nl/meetprogrammas-flora-en-fauna-2022/

# Species
df.species = read.csv('LMF/LMF_Soorten.csv',sep=';', colClasses=c("numeric", "numeric", "factor", "factor")) %>% 
  select(Soortcode, Nednaam, Wetnaam)

# Visits
df.visit <- read.csv('LMF/LMF_kop.csv', sep=';', colClasses=c("factor", "numeric", "character", "numeric", "numeric", 
                                                              "factor", "factor", "factor", "factor", "factor", 
                                                              "factor", "factor", "factor", "factor", 
                                                              "numeric", "numeric", "numeric", "factor", "factor", "factor")) %>% 
  mutate(Datum=as.Date(Datum, format = "%Y%m%d")) %>% 
  select(Plot, Jaar, Dagnummer, Pq_grootte)

# Plots
df.plots <- read.csv('LMF/LMF_plot.csv',sep=';', dec=',', colClasses=c("factor", "numeric", "numeric", 
                                                                       "factor", "factor", "factor", "factor", "factor", 
                                                                       "factor", "factor", "factor", "factor", "numeric", "factor")) %>%
  mutate(X=X*1000, Y=Y*1000) %>%
  #mutate(x = as.numeric(gsub(",", ".", X))*1000,
  #       y = as.numeric(gsub(",", ".", Y))*1000) %>%
  select(Plot, X, Y, Fys_geo_regio)

# Observations, take only latest
df.occurrence <- read.csv('LMF/LMF_tellingen.csv', sep=';', colClasses=c("factor", "numeric", "numeric", "numeric")) %>% 
  select(Plot, Jaar, Soortcode, Bedekking)
df.occurrence <- df.occurrence %>% merge(df.species) %>% group_by(Plot) %>% filter(Jaar == max(Jaar))

# Add plot visit date 
df.extra <- df.occurrence %>% distinct(Plot, Jaar) %>% merge(df.visit)
df.plots <- df.plots %>% merge(df.extra)

# Add abiotic information
df.plots <- df.plots %>% full_join(df.abiotic, by="Plot", suffix = c("", ".abiotic")) %>%
  mutate(X=ifelse(is.na(X), X.abiotic, X),
         Y=ifelse(is.na(Y), Y.abiotic, Y))
```

Wieger's proposed species to analyze

```{r}
# Wamelink, Wieger <wieger.wamelink@wur.nl>
df.n <- df.occurrence %>% group_by(Nednaam) %>% count() %>% arrange(-n)
LMF.dir <- file.path('LMF', '8969892_WageningenEnvironmentalResearch_6708PB_20201006_ww3.xlsx')
species_proposal <- c(read_excel(LMF.dir)$NedNAAM, "Gestreepte witbol", "Struikhei")
species <- (data.frame(Nednaam=species_proposal) %>% merge(df.n) %>% filter(n>=20) %>% select(Nednaam))$Nednaam
chosen.species <- data.frame(Nednaam=species[species != "Struikhei"]) %>% 
  merge(df.species) %>% merge(df.n) %>% select(Nednaam, Wetnaam, n) %>% arrange(-n) 
```

Overlap statistics:

```{r}
c1 <- unique(df.abiotic$Plot)
c2 <- unique(df.visit$Plot)
cx <- intersect(c1, c2)
sprintf("Plots: %d abiotic, %d visit, %d intersection", length(c1), length(c2), length(cx))

c1 <- unique(df.abiotic$DATE)
c2 <- unique(df.visit$Jaar)
cx <- intersect(c1, c2)
sprintf("Years: %d abiotic, %d visit, %d intersection", length(c1), length(c2), length(cx))

c1 <- unique(paste(as.character(df.abiotic$Plot), 'x', as.character(df.abiotic$DATE)))
c2 <- unique(paste(as.character(df.visit$Plot), 'x', as.character(df.visit$Jaar)))
cx <- intersect(c1, c2)
sprintf("(Plot,Year)-pairs: %d abiotic, %d visit, %d intersection", length(c1), length(c2), length(cx))
```

Binary table of each species occurrence at each vegetation plot (0/1)

```{r}
df.y <- df.occurrence %>% select(Plot, Nednaam, Bedekking) %>% mutate(Bedekking=1) %>% spread(Nednaam, Bedekking, fill=0)
```

Take only relevant vegetation plot information and join to species occurrence at each plot

```{r}
df.plots <- df.plots %>% 
  select(Plot, releve_nr, X, Y, Fys_geo_regio, Jaar, Dagnummer, Pq_grootte, `pH-H20`, ORG_STOF, `C/N`, N_totaal, P_totaal, K_CaCl2) %>% 
  mutate(across(`pH-H20`:K_CaCl2, function (x) round(x, digits=2))) %>%
  full_join(df.y[c("Plot", species)])
```

## Add information, spatial joins with geodata

In codes we use X&Y as km to enforce a vague prior

```{r}
NL <- st_read("data_paper/NL.shp")$geometry
NL <- (NL - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
```

Data set of species observations and environmental conditions

```{r}
sf.plots <- df.plots %>% st_as_sf(coords = c("X","Y"), crs=st_crs(NL))
st_geometry(sf.plots) <- (sf.plots$geometry - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
sf.plots[c("X", "Y")] <- sf.plots  %>% st_coordinates()
```

Grid of all possible plots

```{r}
# Grid with rectangles and their centers  #%>% st_as_sf(crs=st_crs(NL))
grid.rect <- st_make_grid(NL,n=100,crs=st_crs(NL), what='polygons', square=T) 
grid.circle <- st_make_grid(NL,n=100,crs=st_crs(NL), what='centers')
index <- which(lengths(st_within(grid.circle, NL)) > 0) #st_intersects(grid.rect, NL)
# Create data frame
sf.grid <- data.frame(grid.circle[index], grid.rect[index]) %>% st_as_sf(crs=st_crs(NL))
sf.grid[c("X", "Y")] <- sf.grid %>% st_geometry %>% st_centroid() %>% st_coordinates()
sf.grid$Plot <- factor(sprintf("Grid%d", 1:nrow(sf.grid)))
```

Add GIS layers: province

```{r}
#https://service.pdok.nl/kadaster/bestuurlijkegebieden/atom/v1_0/bestuurlijke_gebieden.xml
provincie = st_read("BestuurlijkeGebieden_2023.gpkg", "provinciegebied") %>% select(naam)
st_geometry(provincie) <- (provincie$geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)

sf.plots <- sf.plots %>% st_join(provincie , left=T, st_within)
sf.grid <- sf.grid %>% st_join(provincie , left=T, st_within)
#plot(provincie)
```

Add GIS layers: FGR regio

```{r}
# File created by FGR.ipynb
fgr = st_read("FGR_lmf_mapping.gpkg") %>% select(lmf_indeling)
st_geometry(fgr) <- (fgr$geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)

sf.plots <- sf.plots %>% st_join(fgr , left=T, st_within)
sf.grid <- sf.grid %>% st_join(fgr , left=T, st_within)
#plot(fgr)
```

Add GIS layers: soiltype, landuse

```{r}
# https://www.nationaalgeoregister.nl/geonetwork/srv/dut/catalog.search#/metadata/2d3dd6d2-2d2b-4b5f-9e30-86e19ed77a56
# http://www.geodesk.nl/Grondsoorten.htm
## takes ~ 5 mins
landuse <- st_read("bestand_bodemgebruik_2015.gpkg", layer="bestand_bodemgebruik_2015")
soiltype <- st_read("Grondsoortenkaart-2006.shp")
landuse <- landuse %>% transmute(bodemgebruik=relevel(factor(bodemgebruik), ref="Bos"))
soiltype <- soiltype %>% transmute(OMSCHRIJVI=relevel(factor(OMSCHRIJVI), ref="Zand"))
st_geometry(landuse) <- (landuse$geom - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
st_geometry(soiltype) <- (soiltype$geometry - matrix(data=c(155000, 463000), ncol=2)) * matrix(data=c(1/1000, 1/1000), ncol=2)
```

```{r}
## takes ~2 hours
info.plots <- sf.plots %>% select(Plot)
info.plots <- info.plots %>% st_join(landuse , left=T, st_within)
info.plots <- info.plots %>% st_join(soiltype , left=T, st_within)
info.plots <- info.plots %>% as.data.frame %>% distinct(Plot, bodemgebruik, OMSCHRIJVI)
#info.plots %>% saveRDS("plot.info.RDS")
#info.plots <- readRDS("plot.info.RDS")
sf.plots <- sf.plots %>% merge(info.plots, on="Plot")
```

```{r}
## takes ~2 hours
info.grid <- sf.grid %>% select(Plot)
info.grid <- info.grid %>% st_join(landuse , left=T, st_within)
info.grid <- info.grid %>% st_join(soiltype , left=T, st_within)
info.grid <- info.grid %>% as.data.frame %>% distinct(Plot, bodemgebruik, OMSCHRIJVI)
#info.grid %>% saveRDS("grid.info.RDS")
#info.grid <- readRDS("grid.info.RDS")
# sf.grid %>% merge(info.grid, on="Plot") drops the second geometry, so cbind the correctly sorted auxiliary data
sf.grid <- cbind(sf.grid, sf.grid %>% as.data.frame %>% select(Plot) %>% merge(info.grid, on="Plot", sort=F) %>% select(bodemgebruik, OMSCHRIJVI)) 
```

Fill very few NA observations with most common category and combine some categories

```{r}
sf.plots <- sf.plots %>% mutate(
  bodemgebruik= fct_na_value_to_level(fct_collapse(bodemgebruik, 
      Bebouwd=c("Bebouwd exclusief bedrijfsterrein", "Bedrijfsterrein", "Semi-bebouwd", "Glastuinbouw"),
      Spoorterrein=c("Spoorterrein", "Vliegveld")), "Landbouw en overig agrarisch"),
 OMSCHRIJVI=fct_na_value_to_level(relevel(factor(OMSCHRIJVI), ref="Zand"), "Zand"))

sf.grid <- sf.grid %>% mutate(
  bodemgebruik=fct_na_value_to_level(fct_collapse(bodemgebruik, 
      Bebouwd=c("Bebouwd exclusief bedrijfsterrein", "Bedrijfsterrein", "Semi-bebouwd", "Glastuinbouw"),
      Spoorterrein=c("Spoorterrein", "Vliegveld")), "Landbouw en overig agrarisch"),
  OMSCHRIJVI=fct_na_value_to_level(relevel(factor(OMSCHRIJVI), ref="Zand"), "Zand"))
```

Save data

```{r}
chosen.species %>% write.csv(file="data_paper/chosen_species.csv", row.names=F)
sf.plots %>% as.data.frame %>% select(-geometry) %>% write.csv(file="data_paper/data_plots.csv", row.names=F)
sf.grid %>% as.data.frame %>% select(-geometry,-geometry.1) %>% write.csv(file="data_paper/data_grid.csv", row.names=F)
```